---
title: "#TidyTuesday Submission"
subtitle: "Data Science for Biologists, Spring 2020"
author: "Nicholas Paterna"
date: "February 25, 2020"
output:
  html_document:
    highlight: tango
    theme: yeti
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
library(tidyverse)
library(tidylog)
theme_set(theme_bw())

measles <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-25/measles.csv')
```

### TidyTuesday Week 9
> Hi, this is my second time attempting to make a figure for #TidyTuesday! I am a student at Rowan University in Biochemistry with **_ZERO_** experience in coding at all. R is my first language and I am very happy to be learning it, I hope to work in Bioinformatics one day possibly Medical Genomics. I am open to criticism!

```{r TT_Week9, fig.height=4, fig.width=6}

new_england <- c("Maine", "Vermont", "Massachusetts", "Connecticut")
#These are the New England states that presented mmr results in this dataset.
fav_fill <- c("rosybrown3", "lightgoldenrod3", "honeydew3", "royalblue3")
#Here is where I assigned the fills I wanted to a variable.

measles %>%
  select(state, mmr) %>%
  group_by(state) %>%
  filter(state %in% new_england) %>%
  summarize(average_vac = mean(mmr)) %>%
#All of the code above filters out the states of my choosing and then manipulates the data to show only the average vaccination rate per state.
  ggplot(aes(x = fct_reorder(state, average_vac), y = average_vac)) +
  geom_col(aes(fill = state)) +
  scale_fill_manual(values = fav_fill) +
  labs(x = "State",
       y = "Average Vaccination Rate",
       title = "Vaccination Rates in New England States") +
  theme_minimal() +
  theme(legend.position = "none") -> vaccine_plot

ggsave("tidytuesday_week8.png", vaccine_plot, height = 4, width = 6)
vaccine_plot
```

> I know this isn't much of a figure but I wanted to work more on isolating variables in the data that I want to examine. Here I looked at the states with reported data in the New England region, comparing the average vaccination rate per state. I also tried out something my professor showed us in class; to avoid hard coding fill values for multiple plots you can create a variable with the fills you desire. It worked!


