---
title: "Working with strings with `stringr`"
author: "Stephanie J. Spielman"
date: "Data Science for Biologists, Spring 2020"
output:
  xaringan::moon_reader:
    nature:
      highlightLines: true
---
```{css, echo=F}
pre {
  white-space: pre-wrap;
  
}

ul:first-child, ol:first-child {
    margin: 0;
}


.remark-code, .remark-inline-code { 
    color: #326369;
    font-weight: 600;
}
/* Code block code */
.hljs .remark-code-line { 
  font-weight: normal;
}

.pull-left2{
  float: left;
  width: 60%;
}

.pull-right2{
  float: right;
  width: 35%;
}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse=TRUE, fig.height = 4, fig.width = 5)
library(tidyverse)
library(xaringan)
library(patchwork)
```


## Working with strings


```{r, collapse=T}
# https://www.kentuckyderby.com/horses
horse1 <- "Silver Prospector" 
horse2 <- "Candy Tycoon"
horse3 <- "Shoplifted"

all_horses <- c(horse1, horse2, horse3)
print(all_horses)
```

---

## Useful `stringr` functions

side = c("both", "left", "right")
+ General template: `str_ACTION(input_string, ...)`

+ `str_count(input_string, character_to_count)` ==> integer
+ `str_detect(input_string, substring_to_detect)` ==> logical
+ `str_replace(input_string, search, replace)` ==> string
  + `str_replace_all()` replace ALL occurrences instead of only first
+ `str_remove(input_string, substring_to_remove)` ==> string
  + `str_remove_all()` removes ALL occurrences instead of only first
+ `str_starts(input_string, substring)` ==> logical
+ `str_ends(input_string, substring)` ==> logical
+ `str_to_upper(input_string)` ==> string (totally uppercase)
+ `str_to_lower(input_string)` ==> string (totally lowercase)
+ `str_to_title(input_string)` ==> string (totally cap'd after space)
+ `str_trim(input_string, side)` ==> trimmed string 
  + side is one of `c("both", "left", "right")`
+ `str_squish(input_string)` ==> trimmed string 


---

## Counting instances of substrings `str_count()`

```{r}
horse1
all_horses 
```

```{r}
str_count(horse1, "c")
str_count(horse1, "spec")
str_count(all_horses, "c")
str_count(all_horses, "C")
```

---

## Detecting instances of substrings `str_detect()`

```{r}
horse1
all_horses 
```

```{r}
str_detect(horse1, "c")
str_detect(horse1, "spec")
str_detect(all_horses, "c")
str_detect(all_horses, "C")

str_detect(horse1, "Sil", negate=T)
!(str_detect(horse1, "Sil"))
```

---

## Replacing instances of substrings `str_detect()`

```{r}
horse1
horse2
all_horses 
```

```{r}
str_replace(horse1, "c", "!!!")
str_replace(all_horses, "c", "!!!")

str_replace_all(horse2, "y", "WHY")
```

---

## Checking beginnings, endings

```{r}
horse3
all_horses 
```

```{r}
str_starts(horse3, "S")
str_starts(horse3, "s")
str_starts(horse3, "Silver")
str_starts(horse3, "Full")

str_ends(horse3, "definitely not how horse3 ends")

str_starts(all_horses, "Partial")
```

---

## Changing cases

```{r}
horse3
all_horses 
```

```{r}
str_to_upper(horse3)
str_to_lower(horse3)

new_horse <- "mischevious alex"
str_to_title(new_horse)

str_to_upper(all_horses)
```

---

### Trimming whitespace

```{r}

newer_horse <- "Tiz the Law\n\n\n"
str_trim(newer_horse)
```
--
```{r}
newest_horse <- "\r\r\rUntitled"
str_trim(newest_horse)
```
--
```{r}
str_trim(newest_horse, side = "right")
str_trim(newest_horse, side = "left")
```

---

## What is whitespace?

> These can all be used as **regular expressions**


| Symbol | Type of whitespace |
|-------|---------------------|
|`\s`    | any type of whitespace | 
| `\t`   | a tab stroke | 
| `\n`   | a new line (enter on UNIX) | 
| `\r`   | return carriage (enter on PC) |
| `" "`    | I literally typed the space key (but in quotes so you can see there is a space). There's no special symbol, just space!

---

## How to integrate with data analysis?

+ Many ways, but for you guys...
  + `mutate()` new columns based on existing string columns
  + `filter()` rows based on fulfilling certain conditions

```{r}
names(msleep)
msleep %>%
  ## select all columns that are characters (fancy select thing!!!)
  dplyr::select_if(is.character)-> msleep_str

head(msleep_str, 3)
```


---

## Examples!

```{r}
## capitalize genius
msleep_str %>%
  dplyr::mutate(genus_upper = str_to_upper(genus)) %>%
  dplyr::select(genus, genus_upper) %>%
  head(2) ## only show top 3 rows
```
--
```{r}
## title name
msleep_str %>%
  dplyr::mutate(name_titled = str_to_title(name)) %>%
  dplyr::select(name, name_titled) %>%
  head(3)
```

---

```{r}
## any monkeys?!?!??!!!
msleep_str %>%
  dplyr::filter(str_detect(name, "monkey"))
```
--
```{r}
## any rats? trust me this is going somewhere
msleep_str %>%
  dplyr::filter(str_detect(name, "rat"))
```

---

## Introducing REGULAR EXPRESSIONS!

+ `\\b` means "word boundary"

```{r}
msleep_str %>%
  ## Look for pattern: rat must be its OWN WORD
  dplyr::filter(str_detect(name, "\\brat\\b")) #<<
```
---
## One more..

+ `[]` means set of matching characters

```{r}
msleep_str %>%
  ## Now, either case
  dplyr::filter(str_detect(name, "\\b[Rr]at\\b")) #<<
```
