---
title: "TidyTuesday1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(patchwork)
food_consumption <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-18/food_consumption.csv')

write_csv(food_consumption, "food_consumption.csv")

```

## My First Tidy Tuesday

After reading in my data I deciede it would be a good idea to write it as well to keep a .csv saved. The next important step was to actually look at my data

```{r}
food_consumption

```

Already from looking at it 1,430 rows thats alot of information so I began to think of ways to simplify it. 


```{r}
#This chuck will factor my data for use later

food_consumption$food_category<-factor(food_consumption$food_category)


```

## Simiplifying my data

After looking at my data I decieded to simplify it to represent only food categories and using their mean consumption and CO2 emmission. Then I selected the Top 5 CO2 Emmmiters

```{r}

food_consumption %>% 
  group_by(food_category) %>% 
  summarise(consumption=mean(consumption), co2_emmission=mean(co2_emmission)) %>%
  arrange(desc(co2_emmission)) %>%
  top_n(5, co2_emmission)->top_co2_food

top_co2_food

```

After Creating my new tibble I had to create my first plot. This is a bar graph only comparing Food Category to its CO2 Emmision

```{r}
ggplot(top_co2_food, aes(x=food_category, y=co2_emmission, fill=food_category))+
  geom_bar(stat="identity")+theme(legend.position = "none")+labs(x="Food Category",y="Mean CO2 Emmission")->B

B
```

The next plot I created was one comparing the Food Category to the mean Consumption. I did not add the legend or X axis tittle because in my final step I would be combining them. No need for two legends and two X axis if they are going ot be same 

```{r}


  ggplot(top_co2_food, aes(x=food_category, y=consumption, fill=food_category))+
  geom_bar(stat="identity")+theme(legend.position = "none", axis.title.x = element_blank())+labs(y="Mean Consumption")->A
A

  

```

Finally using patchwork I divided the two variables for the graphs I created to compare both bar graphs. Now in this bar graph we see Worldwide mean food consumption comapred to its mean CO2 emmission based on food category

```{r}
A/B->consumption_vs_emmission

ggsave("my_first_Tidy.png", consumption_vs_emmission)

consumption_vs_emmission



```