---
title: "Tidy Tuesday 02/18/2020"
author: "Madeline Greene"
date: "2/20/2020"
output: html_document
---

```{r setup, include=FALSE}

#read in packages and data
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(patchwork)
library(emojifont)


load.emojifont("OpenSansEmoji.ttf")
food_consumption <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-18/food_consumption.csv')



```

## Food Consumption and CO2 Emissions

This is my interpretation of the data that was provided for this TidyTuesday!



```{r, echo = TRUE, fig.height= 3, fig.width= 5}


emojis <- c(emoji("cow"), emoji("egg"), emoji("fish"), emoji("goat"), emoji("milk_glass"), emoji("peanuts"), emoji("pig"), emoji("chicken"), emoji("rice"), emoji("seedling"), emoji("bread")) # tried, but egg, milk, peanut did not load on graph for some reason

food_consumption %>%
  group_by(food_category) %>% #only wanted the food categories
  summarise(mean_co2 = mean(co2_emmission)) %>% #found the average
  ggplot(aes(x = food_category, y = mean_co2, fill = food_category, color = food_category)) + 
  geom_bar(stat = "identity") +
  geom_text(label = emojis, family = "OpenSansEmoji" , size = 10, vjust = -0.3) + # labeled the bar graphs with emojis
  labs( x = "Type of Food", y = "CO2 Emmissions (kg/person/year)", title = "Consumption of Food Types vs. CO2 Emmissions") +
  theme_classic() +
  theme(legend.position = "none") + #took away legend because was big and distracting
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(plot.title = element_text(face= "bold")) -> food_consumption_plot


food_consumption_plot

ggsave("tidy_tuesday_week8.png", food_consumption_plot, height=8, width = 8)


 
```

** Currently in my data science class, we learned dplyr and piping, which I used that in my graph today. My professor mentioned a package for emojis, so I wanted to try that out for my graph. Most of the emojis worked, except three and I am not sure why. I was happy to keep practicing with group_by and summarise, because I was not very confident with that. I also learned about vjust, hjust, and andle for the element_text, which is going to be very useful for future graphs

```



