---
title: "Final Project"
subtitle: "Data Science for Biologists, Fall 2021"
author: "Replace this with your name. It will super awkward if you don't."
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# DON'T CHANGE THE NEXT LINE, except to add NEW options if you have any.
knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo = TRUE, include = TRUE)

# Load all libraries you use HERE:
```

<br><br>

## Introduction




```{r read-examine-data}

```

<br><br>

## Question 1: Modeling

#### REPLACE TEXT WITH YOUR QUESTION. Your question should literally be a level-4 header.

### Methods

<!-- Insert NAMED R chunk(s) here to build your model and make your visualization. --> 

### Answer

Replace this text with the interpretation of your model. If you did a logistic regression, you must explain whether its AUC indicates a strong or weak model. If you did a linear model, you must explain whether BOTH its $R^2$ and RMSE indicate a strong or weak model.


<br><br>

## Question 2

#### REPLACE TEXT WITH YOUR QUESTION. Your question should literally be a level-4 header.

### Methods

<!-- Insert NAMED R chunk(s) here to build your model and make your visualization. --> 


### Answer

Replace this text with your answer to your question.

<br><br>

## Question 3

#### REPLACE TEXT WITH YOUR QUESTION. Your question should literally be a level-4 header.

### Methods

<!-- Insert NAMED R chunk(s) here to build your model and make your visualization. --> 


### Answer

Replace this text with your answer to your question.


<br><br>

## Question 4

#### REPLACE TEXT WITH YOUR QUESTION. Your question should literally be a level-4 header.

### Methods

<!-- Insert NAMED R chunk(s) here to build your model and make your visualization. --> 


### Answer

Replace this text with your answer to your question.
