---
title: "Final Project Proposal"
subtitle: "Data Science for Biologists, Fall 2021"
author: "Your name goes here"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo = TRUE, include = TRUE)
# Load all libraries you use HERE
library(tidyverse)
```



### Introduction

Delete all this text and replace with your introduction of the dataset. *Seriously, it's really awkward if you don't replace this template text! Make sure to always be __KNITTING AND CHECKING THE OUTPUT__ of your Rmds before submitting! Rmd files are the "source code" for the targeted goal of making fancy reports in R in HTML (the knitted stuff!).* Anyways, make sure you also include a properly-formatted link to the data as well. Hint: [this](https://youtu.be/iik25wqIuFo) is a properly formatted link. Knit to see how it looks with your Rmd theme! 

```{r read-and-glimpse-data}
# Code to read in the data:


# Code to dplyr::glimpse() the data:


```

The variables in the dataset are as follows:

<!-- Instructions inside an HTML comment! Use HTML comments to leave yourself notes in the Rmd that you *don't* want to appear in the knitted output. 
Below, OUTSIDE the comment, include a bullet point list or table introducing all variables in the data. 
-->

<br>

### Proposed model

<!-- Below is a "fill in the blanks!" situation! Delete the blanks, and replace, in order, with 1) either: linear or logistic, 2) the name of the variable **in backticks!!**, like `Species`. ---> 

I plan to use a ______________ model to predict the variable ____________.

