---
title: "Materials to incorporate"
author: "Stephanie J. Spielman"
date: "12/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Me previously, factors

#### Understanding factors

Part of this lab/assignment will necessitate working with _factors_, a special type of variable used to work with categories of categorical variables. By default, `readr` assumes that NOTHING is a factor, so we have the full power to determine which variables are factors, and what the _order_ of the categories should be. In R, we call the categories of a factor its _levels_.

> The base R version of the `read_csv` function is `read.csv` and assumes any character is a factor - this problem notoriously causes everyone a lot of major headaches. `readr` rescued us all with its lovely `read_csv` alternative.

The `tidyverse` package [`forcats`](https://forcats.tidyverse.org/) has rich functionality for working with and modifying factors, but for now these features are a bit too advanced (although have at it if you want!), so we will use some convenient base R code.

From the `head(sparrows)` output, we see at least three columns are designed `<chr>`, meaning R thinks those columns are characters - plain old works with nothing interesting linking them. But, we know "better" - it is clear that columns `Sex`, `Age`, and `Survival` are categorical variables with distinct categories (levels!). We can convert these columns to factors like this:

```{r, factor-sex-1}
### Convert Sex to a factor 
sparrows$Sex <- factor(sparrows$Sex)

### If we run head() again, we'll see it's a factor now
head(sparrows)


### Once converted, let's see its levels
levels(sparrows$Sex)


```

**By definition, levels are ordered**, and if no order is specified, R defaults to alphabetical order. To re-order a factor, simply refine the levels with your preferred order of the categories:
```{r, factor-sex-2}
### Make the order Male, Female. 
levels(sparrows$Sex) <- c("Male", "Female")

## Print it out to see
print(levels(sparrows$Sex))
```


This can also all be done at once with the single command below.
```{r, factor-sex-3}
### Convert to factor and order levels as desired at once
sparrows$Sex <- factor(sparrows$Sex, levels=c("Male", "Female"))
```


## Converting an ugly plot to a clean plot

Use the fisheries and instructors datasets to make better figures from the existing.

1. Think about what the GOAL of each figure is. What question is it asking?
2. Create a new compelling figure to answer that question
3. Describe at least 3 things you identified that should be changed for interpretability.

This will involve some amount of tidyr'ing, so this could be a good exercise/hw for having learned tidyr.
This could also be a good option for part of their PROJECT.

## plotting and wrangling: hw-02 from dsbox

The exercise is pokemon-based but seems _mostly_ reasonable as a template for plotting, summary stats, and dplyr

## wrangling FACTORS: hw-03 from dsbox

Again, uses bike rentals dataset but is pretty reasonable for dplyr. Does recoding as FACTORS and some exploratory data analysis, but mostly modeling. We are _not_ ready for modeling guys

## hw-05 from dsbox

**resampling and permutation. we'll want this eventually.**

## PROJECT from ds-box

Very similar to my plan. have some nice _guidelines_ for organization and grading to base off.


## lab-01 is a github lab

## Lab 02: Data wrangling and visualization  

This is a reasonably good joint dplyr and ggplot lab, although we'd need to change the data to be biological.
OR DO WE?