---
title: "Tidy Tuesday"
subtitle: "Data Science for Biologists, Spring 2020"
author: "Julianne Gallagher"
date: "`r Sys.Date()`"
output: 
  html_document:
    highlight: tango
---

<!-- SETUP CODE CHUNK -->
```{r setup, include=FALSE}
## Change the default figure width and height to your liking, but within reason.
knitr::opts_chunk$set(echo = TRUE, fig.width=4, fig.height=3)
library(tidyverse)
library(patchwork)
theme_set(theme_classic())
```



```{r, factor-sex-1}
##load in data
food_consumption <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-18/food_consumption.csv')

food_tibble <- as_tibble(food_consumption)

food_consumption
```




```{r survival-barplot, fig.height= 4, fig.width= 8}
food_consumption %>%
  dplyr::filter(country %in% c("USA", "Brazil")) %>%
ggplot(aes(x=food_category, y=consumption)) + 
geom_col(position= "dodge", aes(fill= country)) + 
scale_fill_manual(values= c("mediumpurple2", "mediumpurple4")) + 
labs(x= "Type of Food", y= "Consumption", title= "Consumption of Food in Brazil vs. USA", fill= "Country") + 
scale_x_discrete(labels= c("Wheat and Wheat Products" = "Wheat", "Nuts inc. Peanut Butter" = "Nuts", "Milk - inc. cheese" = "Dairy")) +
theme_classic() +
theme(plot.title = element_text(face= "bold")) -> tidytues_bar

tidytues_bar

ggsave("first_tidy_tues.png", tidytues_bar, height = 4, width = 8)

```
    
    
**I gained more confidence in working with %in% this #tidytuesday and also reconfirmed some elements in the more practice with ggplot2 assignment. I also learned from this #tidytuesday how to bold the title of my graph. In addition I learned how to abreviate/change the names of labels on the x-axis so they are shorter and do not overlap. By changing the names slightly to make them shorter I was able to make the graph more legable then it previously was.**

