---
title: "The greatest hits of `dplyr` one-table verbs"
author: "Stephanie J. Spielman, PhD"
output: 
  html_document:
    highlight: tango
    theme: flatly
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse=TRUE)
library(tidyverse)
iris <- as_tibble(iris)
```

<br>

+ **Piping syntax** 

  ```{r demo_pipe, eval=FALSE}
  # Without piping
  dplyr_function( data_frame_variable, arguments )
  
  # With piping
  data_frame_variable %>%
    dplyr_function( arguments )
  ```

<br>

+ **[`filter()`](https://dplyr.tidyverse.org/reference/filter.html): Subset _rows_ based on `TRUE`/`FALSE`**

  ```{r demo_filter, exercise=TRUE, exercise.lines = 7}
  # only keep rows where Species == "setosa" is TRUE
  iris %>%
    filter(Species == "setosa")
  
  # only keep rows where the sepal length is greater than 5
  iris %>%
    filter(Sepal.Length > 5)
  ```

<br>


+ **[`mutate()`](https://dplyr.tidyverse.org/reference/mutate.html): Create new _column_**

  ```{r demo_mutate, exercise=TRUE, exercise.lines = 8}
  # create new column called `thiscolumnis5` where the value is 5
  iris %>% 
    mutate(thiscolumnis5 = 5)
  
  # create new column called `sepal_area` as the product of sepal length*width
  iris %>% 
    mutate(sepal_area = Sepal.Length * Sepal.Width)
  ```

<br>


+ **[`distinct()`](https://dplyr.tidyverse.org/reference/distinct.html): Remove duplicate rows**

  ```{r demo_distinct, exercise=TRUE, exercise.lines = 2}
  iris %>% 
    distinct()
  ```

<br>


+ **[`arrange()`](https://dplyr.tidyverse.org/reference/arrange.html): Sort rows according to a given column**

  ```{r demo_arrange, exercise=TRUE, exercise.lines = 6}
  iris %>% 
    arrange(Sepal.Length)
  
  # To arrange in descending order:
  iris %>% 
    arrange(desc(Sepal.Length))
  ```

<br>


+ **[`rename()`](https://dplyr.tidyverse.org/reference/rename.html): Rename columns**

  ```{r demo_rename, exercise=TRUE, exercise.lines = 3}
  iris %>% 
    #       newname     =   oldname    !!!!!!!
    rename(sepal_length = Sepal.Length)
  ```

<br>


+ **[`select()`](https://dplyr.tidyverse.org/reference/select.html): Retain or re-order columns**

  ```{r demo_select, exercise=TRUE, exercise.lines = 8}
  # only keeps those 2 columns
  iris %>% 
    select(Sepal.Length, Sepal.Width)
  
  # remove those two columns and keep the rest
  iris %>% 
    select(-Sepal.Length, -Sepal.Width)
  ```

<br>


+ **[`group_by()`](https://dplyr.tidyverse.org/reference/group_by.html): Create a group for grouped operations**
  + Use `ungroup()` to undo all existing groupings
  
  ```{r demo_group_by, exercise=TRUE, exercise.lines = 10}
  # set up for grouped operations
  iris %>% 
    group_by(Species)
    
  # group and upgroup, as a demonstration:
  iris %>% 
    group_by(Species) %>%
    ungroup()
  ```
  
<br>



+ **[`summarize()`](https://dplyr.tidyverse.org/reference/summarise.html): Create a new data frame that is a _collapsed and summarized_ version**
  
  ```{r demo_summarize, exercise=TRUE, exercise.lines = 10}
  iris %>% 
    summarize(mean_sepal_length = mean(Sepal.Length))
  
  # in combo with group_by!
  iris %>% 
    group_by(Species) %>%
    summarize(mean_sepal_length = mean(Sepal.Length)) %>%
    # to be safe, always ungroup:
    ungroup()
  ```

<br>


+ **[`tally()`](https://dplyr.tidyverse.org/reference/tally.html): Count the number of observations in each group**
  
  ```{r demo_tally, exercise=TRUE, exercise.lines = 5}
  iris %>% 
    group_by(Species) %>% 
    tally() %>%
    # to be safe, always ungroup when you're done
    ungroup()
  ```

<br>


+ **[`count()`](https://dplyr.tidyverse.org/reference/count.html): Handy shortcut to group, tally, ungroup!** 
  
  ```{r demo_count, exercise=TRUE, exercise.lines = 2}
  iris %>% 
    count(Species) 
  ```

