---
title: "How to set up your GitHub Token in RStudio Cloud"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=FALSE)
```
<br><br>

### The video: [YouTube Link](https://youtu.be/h1ykWSTbUuI)

<br><br>

### Commands you'll need to use:

<br>

#### Command 1: Set your name and email

```{r}
# replace "Your Name" with your literal name (keep the quotes!!!)
# replace "your@email.com" with the email you used to sign up for github.com (keep the quotes!!!)
usethis::use_git_config(user.name = "Your Name", user.email = "your@email.com")
```
<br><br>

#### Command 2: Create Github Token

```{r}
# will open github.com for you to make a token
usethis::create_github_token() 
```

<br><br>

#### Command 3: Tell RStudio Cloud about that token

```{r}
# will pop up a window where you PASTE the secret token!
credentials::set_github_pat()
```
<br><br>

#### Command 4: Make sure it worked!

> You should copy/paste the output from this command from the R Console into the Canvas assignment "Proof of RStudio Cloud/Github Setup." Submitting this copy/pasted text fully completes the assignment!! Completing this on time is part of your HW2 grade.

```{r}
usethis::git_sitrep()
```