---
title: "EXAMPLE PROJECT"
subtitle: "Data Science for Biologists, Fall 2020"
author: "Stephanie J. Spielman"
output: 
  html_document:
    theme: simplex
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo = TRUE, include = TRUE)
# Load all libraries you use HERE
library(tidyverse)
library(broom)
```

## Introduction

In this project, I analyze a [dataset of coffee ratings](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-07-07/readme.md) from the `#tidytuesday` repository. This dataset contains information about coffees and their ratings from around the world:

```{r read-data}
coffee_ratings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-07/coffee_ratings.csv')

coffee_ratings
```


## Question 1

### Do Arabica or Robusta coffees have a higher mean flavor value?

#### Methods

```{r question1, fig.width = 4, fig.height=3}
# calculate mean flavors for each species
coffee_ratings %>%
  group_by(species) %>%
  summarize(mean_flavor = mean(flavor))

# visualize flavor distributions and means per species
ggplot(coffee_ratings, aes(x = species, y = flavor, fill = species)) + 
  geom_violin(alpha = 0.8) + 
  scale_fill_manual(values=c("purple", "chocolate4")) + 
  theme_bw() + 
  stat_summary()
```

#### Answer
Robusta has a slightly higher mean flavor at 7.63 compared to Arabica's mean flavor of 7.51.



## Question 2

### To what extent does a coffee's processing method explain its flavor rating?

#### Methods

First, I prepare the data to model.
```{r question2-prep}
# prepare data to model
coffee_ratings %>%
  select(flavor, processing_method) %>%
  drop_na() -> coffee_ratings_model_data

# display data
coffee_ratings_model_data
```

Then, I assess the assumption of equal variance across groups. The "Washed/Wet" category has a larger spread compared to the other distributions, but the difference is not severe. 
```{r question2-assumptions, fig.width = 4, fig.height=3}
# Assess assumption of roughly equal variance among processing_method groups
ggplot(coffee_ratings_model_data, aes(x = processing_method, y = flavor)) + geom_point()
```

Now, I build the model and assess the assumption of normal residuals. In the QQ plot, most points fall along the line, so the residuals are indeed normally distributed.

```{r question2-buildmodel, fig.width = 4, fig.height=3}
# build model
fit <- lm(flavor ~ processing_method, data = coffee_ratings_model_data)

# Assess assumption of normal residuals
qqnorm(fit$residuals)
qqline(fit$residuals)
```

Now, I examine model output:

```{r question2-examinemodel}
summary(fit)
```

Finally, I make a figure to accompany my model:

```{r question2-plotmodel, fig.width = 7, fig.height=4}
ggplot(coffee_ratings_model_data, aes(x = fct_reorder(processing_method,flavor), y = flavor)) + 
  # reorder x axis based on flavor
  geom_boxplot(aes(fill = fct_reorder(processing_method,flavor))) + 
  scale_fill_brewer(palette = "Dark2") + 
  xlab("Processing method") + ylab("Coffee flavor") + 
  annotate("text", x = 1.2, y = 9, label = "R^2 == 0.01289", size=4,  parse=TRUE)  +
  theme(legend.position = "none", 
        # smaller font to fit on axis
        axis.text.x = element_text(size=7.5))
```


#### Answer
This regression yielded an adjusted $R^2 = 0.01289$ with a significant P-value of 0.0007458. These results suggest that processing method, while significant, is not a strong predictor of coffee flavor, but is actually a very weak predictor. In other words, there is definitely some kind of relationship between processing method and flavor, but the relationship is extremely weak.



