---
title: "Homework #10: Due Tuesday 4/13/21 to Canvas by 1 pm"
subtitle: "Data Science for Biologists, Spring 2021"
author: "Replace this with your name"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo=TRUE, include=TRUE, eval=TRUE)
library(tidyverse)
library(broom)
library(modelr)

# Uncomment and define this variable as your Banner ID number. You must use this variable in Part 2 to set your seed.
#my_banner_id <- 
```


> Caution: This homework is relatively short, which means each question is worth a lot of points. This is both a blessing and a curse. Plan accordingly.


## Instructions

For your assignment, you will be building and evaluating a linear model. Complete your assignment within this template. Before submission, you may delete instructions if you want, but it is not required. 


As part of filling in this template, you must specify in the YAML front matter...

+ Your name should go in the "author" field
+ An *Rmarkdown document* theme of your choosing should be indicated
  + Built-in options: [Gallery of themes](https://www.datadreaming.org/post/r-markdown-theme-gallery/)
  + If you want to scour the internet for other themes in other packages, please feel free! That library will have to be loaded in your setup R chunk, *but NOT the code to install!!!* You will likely have to install the library yourself in your R homework project (Console!), and as needed, I will install it for myself when grading to ensure your code works.
+ A *code highlighting* theme of your choosing should be indicated: [Gallery of code highlighting](https://www.garrickadenbuie.com/blog/pandoc-syntax-highlighting-examples/)
  + Ignore all code in this blog post - it is useful only to see how the different syntax highlighting themes appear so you can find your favorite
+ You are allowed to further modify the template *only to add line spacing with* `<br>` per your personal preference. Everything else should remain as templated.
  + Helping you understand your own code
  + Increasing the transparency of the data analysis/visualization process for your _reproducible code report_. Your comments assist viewers
  + **You must ensure that your comments make sense for the code they are referring to, or else there may be deductions.** For example, the comment `#builds a histogram()` should not be associated with, for instance, `scale_fill_manual()` code, which does not build histograms.



### Please read these additional instructions carefully:

+ You must define your Banner ID variable in the setup chunk and use this quantity to set your random seed in Part 2.
+ All of your code and output (including plots!) must be displayed in the knitted HTML output. **Do not modify any chunk parameters, except `fig.width` and `fig.height`, as needed.** It will likely be needed to make your figures appear at an appropriate size in the final knitted HTML.
+ **ALWAYS RUN AND CHECK CODE ONE LINE AT A TIME!!!**
+ Unless otherwise stated, your wrangling code to answer a given question **MUST** all be part of the same connected pipeline with `%>%`. 
+ You will not receive any credit for answers without corresponding code that produces those answers.
+ Make sure your code runs WITHOUT an errors before submitting. *A 10% penalty is automatically applied for submissions with bugs that lead to errors.*
  + If there are errors in your code, you should *comment out* the code so that it does not actually run. This is **BETTER** than allowing buggy code to run - it shows me you attempted the code, but understood that it didn't work properly. If you leave buggy code in the script without commenting it out, there will be deductions.
+ Be careful not to *duplicate* your code. For example, do not write code on one line, and then on the next line write it again and save it to a variable. Only include the second line in the final version of your script!
+ **As always, you are encouraged to work together and use the class Slack to help each other out, but you must submit YOUR OWN CODE.** 

<br><br>

<!-- You can delete instructions above if you want, but it won't matter for your grade either way. Just be sure to not delete anything below. -->


## Questions


This assignment will use an dataset of various physical measurements from 250 adult males. Your goal for this assignment is to build and evaluate a model from this data to **predict body fat percentage** (column `Percent`) in adult males, and then use this model to predict future outcomes. Age is measured in years, weight in pounds, height in inches, and all other measurements are circumference measured in cm.

```{r read_data, collapse=T}
bodyfat <- read_csv("https://raw.githubusercontent.com/sjspielman/datascience_for_biologists/master/data/bodyfat.csv")

glimpse(bodyfat)
```

### Part 1: Build a model using stepwise model selection


1. Using the `step()` function, determine the most appropriate model to explain variation in bodyfat percentage in this data, and reveal the full model output with the function `summary()` and/or `tidy()`. 
*After the code*, write a properly-formatted (make sure to knit and check!) bullet point list stating the final predictors. You do not need to include information about the specific coefficient values or P-values - just bullet out what the predictors are.

```{r perform_model_selection}


```



2. Use both the `rsquare()` and `rmse()` functions from the `modelr` package to reveal the model's $R^2$ and RMSE. In addition to commenting your code, below the code chunk, explain **whether and _why_** these values suggest the model is highly explanatory and likely has low error, or whether the model is weakly explanatory and likely has higher error. Your answer must specifically reference the calculated $R^2$ and RMSE values.


```{r rsquare_rmse}


```

  **Replace this text with your answer, but make sure your answer is still bolded.**
    
    

<br><br>

### Part 2: Evaluate the model 

1. In the code code below, create a testing-training split where the training data comprises a random *70%* subset of the total `bodyfat` dataset. Build the model using the training data, and then evaluate how the model performs on both the training and testing splits. After the code, provide a brief explanation (1-2 sentences) about whether the model is likely overfit to the training data or not, and these results suggest the model will do a good or poor job predicting future outcomes.

    Your code must have the following components:

    + First, set your seed AS YOUR PERSONAL BANNER ID. YOU MUST USE YOUR BANNER ID AS THE NUMBER, which you should have defined in the `setup` code chunk!!! **If you do not do this properly, there will be unpredictable mismatches between your written answer and the result your code produces when run. You must set your seed, and use it consistently.**
    + Create a variable to represent the 70% split instead of hardcoding this value.
    + There may be other code you need to run to help you interpret RMSE. This code should be included in the code chunk as part of the homework (i.e., not run separately in Console.) 

```{r train_test_code}


```

    **Replace this text with your answer, but make sure your answer is still bolded.**

2. Make either a faceted barplot or a lollipop plot to simply show your $R^2$ and RMSE results for a little wrangling fun. Your plot should follow these guidelines:

    + Facet by measurement ($R^2$ and RMSE). Each of the two plot panel should show one of those quantities
    + The x-axis should be the dataset (testing or training).
    + The height of the bar or lollipop should be the given quantity ($R^2$ and RMSE)
    + Style in any way of your choosing, adhering to professional best-practices we have discussed. You can use the term "R-squared" instead of fighting too hard to get the plot to have superscript, too (but figure out how if you want!).
    + In order to create this plot, you will need to first create the tibble that will contain your data to plot. See [Chapter 10](https://r4ds.had.co.nz/tibbles.html) of *R for Data Science* to learn a few strategies for making tibbles.

```{r rsq_rmse_plot}


```

<br><br>

### Part 3: Predictions

A new man has arrived, and we want to use our model to predict his body fat percentage. Predict what the body fat percentage will be for a man with the following physical attributes. Your code should reveal the predicted bodyfat percentage.

**Hint: Do you need all this information? In fact you do not! For full credit, make sure your code is only considering body measurements that are relevant for answering the question.**


+ Age:      34
+ Weight:   172.2
+ Height:   64.25
+ Neck:     33.4
+ Chest:    101.4
+ Abdomen:  91.3
+ Hip:      104.7
+ Thigh:    62.5
+ Knee:     38.1
+ Ankle:    26.5
+ Biceps:   33.2
+ Forearm:  31.4
+ Wrist:    17.9

```{r predict_percent}


```

<br><br>

### Optional Bonus

Make a _faceted scatterplot_ from the full `bodyfat` dataset showing the relationship between `Percent`, and each of the model's predictors separately. In other words, the Y-axis should be `Percent`, the X-axes should each show a different predictor along the faceting. You can style this plot however you choose so long as it adheres to best practices. To accomplish this plot, you will need to *pivot* the data in someway! 


```{r bonus_plot}

```