---
title: "HW5 Instructions"
subtitle: "Data Science for Biologists, Spring 2020"
author: "R script due Thursday 10/16/20 by 12:01 PM to Canvas"
output: 
  html_document:
    highlight: tango
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
library(tidyverse)
theme_set(theme_classic())
```


## Instructions

For your assignment, you will be honing your `dplyr` and `ggplot2` skills using a previous Tidy Tuesday dataset about meteorites. The dataset is described and documented [here](https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-06-11).

Please read these additional instructions carefully:

+ CODE MUST BE PRESENT FOR READING FILES AND SAVING PLOTS. NO CREDIT WILL BE GIVEN FOR ANY PLOT WHERE YOU DO NOT HAVE ASSOCIATED CODE TO PERFORM ALL TASKS.
  + NEVER CLICK THE "IMPORT DATASET" BUTTON. YOU MUST USE `read_csv()`.
  + NEVER CLICK THE "EXPORT" BUTTON IN THE "Plots" PANE. YOU MUST USE `ggsave()`.
+ All plots must be professionally labeled with clean X and Y axes. You are not required to include a title, subtitle, or caption, but you may include one or more of these labels if you choose.
+ All plots using aesthetic color/fill mappings _MUST USE_ a non-default color scale. Similarly, all legend titles must be professionally re-named from their default.
+ All plots must be clearly legible with an appropriate aspect ratio. This means you may need to change the output plot size using arguments `width` and `height` when calling `ggsave()`. *Make extra sure to LOOK AT YOUR SAVED PLOTS* and tweak the size until it looks clearly professional.
  + You may also need to modify certain theme components to achieve a professional look!
+ Ask me for help AT LEAST 24 hours before the deadline. I cannot guarantee time to help you if only short notice is given. Given my schedule, I can probably (unfortunately) guarantee that I won't have time to help.
+ Write all code yourself and include COMMENTS for each plot you create.
+ Please *comment out* (add `#` in front of) code that causes errors or bugs. This will ensure that your code runs without errors, while also showing me your code attempt if you only got part-way through recreating the plot.

## Questions

> All questions should be answered in provided HW R script template. 

```{r}
met %>%
  filter(year >= 2008, class == "LL5") %>%
  tally()
```


```{r}
met %>%
  filter(name_class == "Valid") %>%
  group_by(class) %>% 
  # mass in kg
  summarize(mean_mass = mean(mass/1000)) %>% 
  arrange(desc(mean_mass))
```


```{r}
met %>%
  drop_na(year) %>%
  mutate(century = case_when(year < 1900~ "1800s",
                             year >= 1900 & year < 2000~ "1900s",
                             year >=2000 ~"2000s")) %>%
  count(century, name_type)
```

```{r}
# New jersey rests between 39-41.5 lat and -74 - -75 long. 
met %>%
  filter(lat >=39, lat <=41.5, long <= -74, long >= -75)
```



```{r}
# find all the found meteorites that start with the letter of your name and fell in your birthyear
met %>%
  filter(fall == "Found") %>%
  filter(str_starts(name, "S"), year == 1988)
```



```{r}
met %>%
  count(class) %>%
  arrange(desc(n))
```

```{r}
met %>%
  filter(str_starts(class, "Iron"), year >= 1900, year < 2000) %>%
  group_by(class) %>%
  summarize(min_mass = min(mass, na.rm=T)) %>% 
  arrange(min_mass)
```

```{r}
met %>%
  filter(str_starts(class, "Iron"), year >= 1900, year < 2000) %>%
  group_by(class) %>%
  summarize(min_mass = min(mass, na.rm=T)) %>% 
  arrange(min_mass)
```

```{r}
met %>%
  drop_na(year) %>%
  filter(str_starts(class, "Iron")) %>%
  mutate(century = case_when(year < 1900~ "1800s",
                             year >= 1900 & year < 2000~ "1900s",
                             year >=2000 ~"2000s")) %>%
  count(class, century) %>%
  ggplot(aes(y = fct_reorder(class,n), x = n)) + geom_col() + facet_wrap(~century)
```




