---
title: "Tidy Tuesday #2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)


measles <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-25/measles.csv')

```

## First Step Look at the Data



```{r}
measles
```

## Wrangle the data

Then wrangle the data into the what I truly want to compare. In this case I want to compare the average MMR vaccine rate between private schools and public schools according to school year.

```{r}
measles %>%
  select (year,type, mmr) %>%
  drop_na()%>%
  filter(type %in% c("Public","Private")) %>%
  group_by(year,type) %>%
  summarise(mmr=mean(mmr)) %>%
  rename(average_mmr=mmr)-> average_mmr_school

average_mmr_school
```
## Plot the Data

Once everything as been wrangled, I displayed the data in a way that is color-blind friendly. In this graph it can be seen that over the a single school year the MMR vaccinae rate has dropped in private schools across the nation

```{r}
ggplot(average_mmr_school, aes(x=factor(year), y=average_mmr, fill=type))+
  geom_bar(stat="identity", position="dodge")+
  labs(x="School Year", y= "Percentage of Average MMR Vaccination \n Rate ", title = "Percentage National MMR Vaccination Rate \n per School Year In the United States",fill= "School Type")+
  scale_fill_manual(values=c("black","grey"))->Tidy_tuesday_2

Tidy_tuesday_2
```

```{r}
ggsave("tidy_tuesday.png", Tidy_tuesday_2)
```
