---
title: "Exploratory analysis redux"
output: 
  html_document:
    theme: readable
    highlight: pygments
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 6, fig.height = 4, message=F, warning=F)
library(tidyverse)
library(ds4b.materials)
library(ggforce)
```


## Punchlines:

+ Read your words out loud to check that they make sense and use the spell check.
+ Questions should be about one or more variables in the data.
+ Be _specific_ and _precise_ about what are you asking about.
+ Questions should not be very creative. Don't overthink - keep it simple!
  + "Reverse engineering" questions from creative plots can lead to questions that don't make sense.
+ Limit questions to the scope of the data. Being too fancy = being wrong.
+ *Do not even think about cause and effect unless the data was an experiment evaluating cause and effect.*

## Good (non-creative!) places to explore

+ Distributions of a numeric variable
  + Shape and skew, mean/median, range, spread
+ Counts of a categorical variable
+ Relationships between variables. *Seriously, never cause/effect.*
  + Relationship between two numeric variables - positive? negative? strong? weak?
  + Relationship between two categorical variables - joint counts!
  + Relationship between numeric and categorical - "how do means compare?" etc.


## Peer review of student exploratory analyses

<br><br>


#### How many bins are in the histogram?

There are 14 bins in the histogram.

```{r}
ggplot(wine) + 
  aes(x = color) + 
  geom_density(color = "black", 
                 fill = "red") + 
  labs(x = "Color intensity of the wine", 
       y = "Count")
```






<br>


####  What is the difference between the distributions of wing size between sexes?

Females wing size tends to be larger than males, on average.  

```{r}
ggplot(damselfly) +
  aes(x = Wing.size, fill = Sex) + 
  geom_density(alpha = 0.6) +
  labs(x = " Wing size ") +
  theme_classic()
```



<br>

#### Is the distribution of urea unimodal, bimodal, or multimodal in this histogram?
This histogram is multimodal because it has only one peak. 

```{r}
ggplot(urine) +
  aes(x = urea) +
  geom_histogram(fill = "orange1",
                 color = "black",
                 size = 0.5,
                 bins = 25) +
  labs(x = "Urea concentration",
       y = "count")

```


<br>


#### Is there a relationship between osmolarity of urine and specific gravity of urine?

Yes there exists a positive relationship between the two. As osmolarity increases, specific gravity increases as well.

  
```{r}
ggplot(urine) +
  aes(x = osmo,
      y = gravity) +
  geom_point(size = 3) +
  geom_smooth(method = "lm") + 
  labs(x = "Osmolarity",
       y = "Specific Gravity") +
  theme_bw() 
```


<br>


#### What is the relationship between the body mass and age of women, for whether or not they have diabetes?

There is no relationship overall because the trend line is flat.

```{r}

ggplot(pima) +
  aes(x = age,
      y = bmi) + # added color to aes to map color to diabetic
  geom_point() +
  scale_color_manual(values = c("darkorange2",
                               "black")) +
    geom_smooth() + 
  labs(x = "age",
       y = "BMI")
```


<br>


#### Is a person with a higher urea concentration more likely to have kidney stones?

As urea concentration increase the likelihood of crystal formation of kidney stones is more likely to occur as seen as a higher density of those with crystals have a high concentration of urea.
  

```{r}
ggplot(urine) +
  aes(x = urea,
      fill = crystal) +
  geom_density(color = "black",
               alpha = 0.50) +
  scale_fill_brewer(palette = "PuRd") +
  theme_classic() +
  labs(x = "Urea Concentration", y = "Density")
```


<br>


#### Is there a BMI that women did not have have any pregnancies?

Yes the BMI range of 60-67 the women did not have any pregnancies.


```{r}
ggplot(pima) + 
  aes(x = bmi) + 
  geom_histogram(fill = "yellow",
                 color = "chocolate3", bins = 15) + 
  labs(x = "BMI", #labels
       y = "Number of times a Women is Pregnant", 
       title = "Number of times a Women is Pregnant and their BMI's")
```


<br>



#### What is the higest insulin vaule in people without diabetes ?

The highest insulin value for people with out diabetes is around 750.

```{r}
pima %>%
  filter(diabetic ==  "No") %>%
  summarize(max_insulin = max(insulin))

ggplot(pima)+
  aes(x=diabetic,
      y=insulin)+
  geom_boxplot() + 
  geom_sina()+
  theme_bw()+
  labs(x="Diabetes",
       y="Insulin (μU/mL)")
```


<br>

#### Is there sexual selection for wing size?

Females choose their mate based on wing size, with the average male having a wing size of 1900.

```{r}
ggplot(damselfly) +
  aes(x = Wing.size, fill = Sex) +
   scale_fill_brewer(palette = "Set3") +
  geom_density(alpha = 0.8) +
  facet_grid(vars(Mating.status)) +
  theme_minimal() +
  labs(x = "Wing size",
       title = "Wing Size in Mating Pairs") 
```

<br>


#### Does abdomen length play a role in damsel fly sexual selection?

Abdomen length seems to play a role in damsel fly sexual selection, as those with smaller abdomens seem to be more likely to be mated. This was not what I actually expected.

```{r}
ggplot(damselfly)+
  aes(fill=Mating.status, y=Abdomen.length, x=Sex)+
  geom_violin()+
  labs(x="Fly sex", 
       y="Abdomen Length", 
       fill="Mating Status")+
  scale_fill_brewer(palette="Pastel1")+
  theme_light()

```
