---
title: 'Template: Homework #9'
author: "Data Science for Biologists, Fall 2021"
date: "Submit this completed Rmd file to Canvas on Wednesday 11/10/21 by 11:59 PM"
output: 
  html_document:
    theme: lumen
    highlight: tango
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 6, fig.height = 4)

# Always load libraries in the setup chunk:
library(tidyverse) # allows you to use ggplot2, readr, forcats, etc. 
library(ds4b.materials) # contains all but the coffee dataset

# Read in the coffee dataset, used in part 4:
# This code also gets rid of the "bad" point!
read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-07/coffee_ratings.csv')  %>%
  filter(total_cup_points != 0)  -> coffee_ratings

```



## Part 1

<br>

#### Question 1.1

```{r q1.1}


```


#### Question 1.2

```{r q1.2}


```



#### Question 1.3

```{r q1.3}


```



#### Question 1.4

```{r q1.4}


```



#### Question 1.5

```{r q1.5}


```



#### Question 1.6

```{r q1.6}


```



#### Question 1.7

```{r q1.7}


```



#### Question 1.8

```{r q1.8}


```

<br><br>

## Part 2
<br>

#### Question 2.1

```{r q2.1}


```

<br> 

#### Question 2.2

```{r q2.2}


```

<br> 

#### Question 2.3

```{r q2.3}


```

<br>

#### Question 2.4

**Replace this bold text above the chunk with your question.**

Replace this plain text above the chunk with your answer.


```{r q2.4}
coffee_ratings %>%
  # add the coffee_id column
  mutate(coffee_id = 1:n())

```


<br><br>

## Part 3
<br>

#### Question 3.1

Replace this plain text above the chunk with your answer to the question.

```{r q3.1}


```

<br> 

#### Question 3.2

Replace this plain text above the chunk with your answer to the question.

```{r q3.2}


```





<br><br>


## Part 4
<br>

#### Question 4.1

```{r q4.1}


```

<br> 

#### Question 4.2

```{r q4.2}


```