---
title: "HW3 KEY"
subtitle: "Data Science for Biologists, Spring 2020"
#date: "12/28/2019"
output: 
  html_document:
    highlight: tango
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, fig.width=6, fig.height=4)
library(tidyverse)

readr::read_csv("sparrows.csv") %>%
  dplyr::mutate(Age = as.factor(Age), 
                Survival = as.factor(Survival), 
                Sex = as.factor(Sex)) -> sparrows
```


## How to load the data

For each session, copy and paste this code into *Console*:
```
library(dplyr)
readr::read_csv("sparrows.csv") %>% 
  dplyr::mutate(Age = as.factor(Age), 
                Survival = as.factor(Survival), 
                Sex = as.factor(Sex)) -> sparrows
```

### Plot 1
<br>
```{r}
ggplot(sparrows, aes(x = Skull_Length, y = Skull_Width)) + 
  geom_point(fill = "blueviolet", color = "black", pch=22, size=3) + 
  labs(x = "Sparrow skull length (mm)",
       y = "Sparrow skull width (mm)", 
       title = "Plot 1") + 
  geom_smooth(method = "lm", se=FALSE, color = "black") + 
  theme_bw()
```

<br><br>


### Plot 2
<br>
```{r}
ggplot(sparrows, aes(x = Sternum_Length)) + 
  geom_histogram(fill = "bisque3", color = "chocolate4", bins = 12) + 
  labs(x = "Sparrow sternum length (mm)",
       y = "Number of sparrows", 
       title = "Plot 2") + 
  theme_bw()
```


<br><br>

### Plot 3
<br>
```{r}
sparrows$Sex <- factor(sparrows$Sex, levels = c("Male", "Female"))
ggplot(sparrows, aes(x = Sternum_Length, fill = Sex)) + 
  geom_density(alpha = 0.5) +
    labs(x = "Sparrow sternum length (mm)",
       y = "Density", 
       title = "Plot 3") + 
  theme_classic() +
  theme(legend.position = "bottom")
```

Hints: 

+ Make sure your *order of variables* agrees with this plot!! Think: *factors*. 
  + Bear in mind that if you change any data here, it will remain changed throughout the rest of your script. Maybe later you'll need to have code to change things back??
+ This plot places the legend at the bottom of the plot. This is actually a modification to the plot's theme. Because `ggplot` plots are just added components *on top of one another*, you will need to change the legend position AFTER you set a baseline theme (think about why this MUST BE THE CASE, and experiment with it!). The legend can be repositioned with the following code:
`theme(legend.position = "bottom")`. To learn more (but not to submit!), you can see what happens if you use "top", "left", or "right" instead of "bottom".

<br><br>


### Plot 4
<br>
```{r}
sparrows$Age <- factor(sparrows$Age, levels = c("Young", "Adult"))
ggplot(sparrows, aes(x = Age)) + 
  geom_bar(color = "darkgoldenrod", fill = "grey30", size=4) +
    labs(x = "Sparrow age",
       y = "Nummer of sparrows", 
       title = "Plot 4") 
```

<br><br>

### Plot 5
<br>
```{r}
ggplot(sparrows, aes(x = Age, fill = Survival)) + 
  geom_bar(position = position_dodge(), color = "black") +
    labs(x = "Sparrow age",
       y = "Nummer of sparrows", 
       title = "Plot 5") + 
  theme_minimal()
```

<br><br>

### Plot 6
<br>
```{r}
ggplot(sparrows, aes(x = Survival, y = Weight, color = Weight)) + 
  geom_jitter(pch = 18, size=3, width = 0.1, alpha = 0.4) +
    labs(x = "Sparrow survival status",
       y = "Sparrow weight (g)", 
       title = "Plot 6") + 
  stat_summary() +
  theme_classic() + 
  theme(legend.position = "top")
```

Hints:

+ See the *black point with small lines* in each distribution? This represents the *mean and standard error (SE)* of each distribution. In this case, the standard errors are very small, so the lines are a bit tricky to see, but they are there! These points are conveniently, amazingly, and automatically added with the plot component, `stat_summary()`. This is a special `ggplot2` function which can easily add a *summary statistic* onto a plot. No arguments are necessary - by default, it plots mean and SE.
  + It will reveal a warning message: "No summary function supplied, defaulting to `mean_se()`," which means you used it CORRECTLY!!

<br><br>

### Plot 7
<br>
```{r}
ggplot(sparrows, aes(x = Tarsus_Length, y = Humerus_Length, color = Weight, shape = Survival)) + 
  geom_point(size = 2, alpha = 0.7)+
    labs(x = "Tarsus Length (mm)",
       y = "Humerus Length (mm)",
       title = "Plot 7") +
  theme_classic()
```

<br><br>

## BONUS PLOT
<br> 
*Including this plot is optional, but HIGHLY encouraged. Partial credit is always given for attempts, but as always, there should still not be any bugs.*

<br><br>

```{r}
ggplot(sparrows, aes(x = Age, color = Sex, y = Weight)) + 
  geom_violin(alpha = 0) +
  geom_point(pch = 21, position = position_jitterdodge(jitter.width = 0.3), color = "black", aes(fill = Sex)) +
    labs(x = "Sparrow age",
       y = "Weight (g)",
       title = "BONUS PLOT") +
    theme(legend.position = "bottom")
```

Hints:

+ There are *two* geoms used here, one on top of the other. Since `ggplot2` adds plot layers in order on top of one another, think carefully about which geom comes first in the code.
+ The geoms use different but similar mappings. *Think carefully about where these aesthestics need to be placed in the code?* For SHARED mappings, place in the `aes()` inside the main `ggplot()` call. For mapping UNIQUE to a geom, place the `aes()` inside the relevant geom itself.
+ Having no fill is like having total transparency.
