---
title: "Data Science for Biologists Resources"
author: "Stephanie J. Spielman, PhD"
output: 
  html_document:
    highlight: tango
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse=TRUE)
library(tidyverse)
iris <- as_tibble(iris)
```


## General resources

+ [Twitter for R programmers](https://www.t4rstats.com/)
  + Geared towards programmers but serves as a great general "welcome to the Twitter #rstats community"! 
+ [Fundamentals of Data Visualization](https://clauswilke.com/dataviz) by Claus O. Wilke
+ [R for Data Science](https://r4ds.had.co.nz/) by Garrett Grolemund and Hadley Wickham
+ [Data Visualization: A practical Introduction](https://socviz.co/) by Kieran Healy
+ [`tidyverse` homepage, official documentation, and cheatsheets](https://www.tidyverse.org/)
+ [STAT545 Resources (**EXCELLENT** background)](https://stat545.com/)
+ [Introduction to Data Science e-book](https://rafalab.github.io/dsbook/)
+ [RStudio Resources (all levels of videos/tutorials)](https://resources.rstudio.com/)
+ [Calling Bullshit Course Website](https://callingbullshit.org/) by Carl Bergstrom and Jevin West

## Materials from Dr. Spielman
+ [Introduction to R tutorial](http://htmlpreview.github.io/?https://github.com/sjspielman/datascience_for_biologists/blob/master/fall2020/tutorials/introduction_to_R.html)
+ [Introduction to R exercises](https://rowanbiosci.shinyapps.io/intro_R/)
+ [Introduction to `ggplot2` exercises](https://rowanbiosci.shinyapps.io/intro_ggplot2/)
+ [Introduction to `dplyr` exercises](https://rowanbiosci.shinyapps.io/intro_dplyr/)

## `ggplot2` resources

1. [The official ggplot2 documentation](https://ggplot2.tidyverse.org/reference/) provides information for using `ggplot2` functions and commands
2. [The ggplot2 CheatSheet](https://github.com/rstudio/cheatsheets/blob/master/data-visualization-2.1.pdf) is an _information rich_ overview of much of `ggplot2`'s functionality
3. [Chapter Five from "Fundamentals of Data Visualization"](https://serialmentor.com/dataviz/directory-of-visualizations.html) provides a "directory" of different types of plots which exist for different types of data
4. [Chapter Three from "Data Visualization: A practical introduction"](https://socviz.co/makeplot.html) introduces `ggplot2` usage
5. [Chapter Three from "R for Data Science"](https://r4ds.had.co.nz/data-visualisation.html) introduces `ggplot2` usage
6. [All the R colors you ever wanted to use!!](https://github.com/sjspielman/datascience_for_biologists/raw/master/spring2020/rmd_lessons/static/Rcolor.pdf)
7. [Reference: Customizing Colors and Fills](http://htmlpreview.github.io/?https://github.com/sjspielman/datascience_for_biologists/blob/master/fall2020/tutorials/color_fill_scales.html)
7. [Reference: Customizing themes](http://htmlpreview.github.io/?https://github.com/sjspielman/datascience_for_biologists/blob/master/fall2020/tutorials/themes.html)
7. [Reference: Customizing axes](http://htmlpreview.github.io/?https://github.com/sjspielman/datascience_for_biologists/blob/master/fall2020/tutorials/axes.html)


## `dplyr` resources


+ [The official `dplyr` documentation](https://dplyr.tidyverse.org/reference/) provides information for using `dplyr` functions and commands
+ [The `dplyr` CheatSheet](https://github.com/rstudio/cheatsheets/blob/master/data-transformation.pdf) is an _information rich_ overview of much of `dplyr`'s functionality
+ [The `dplyr` vignette](https://dplyr.tidyverse.org/articles/dplyr.html) introduces most of the fundamental `dplyr` functionality
+ [Chapter Five from "R for Data Science"](https://r4ds.had.co.nz/transform.html) introduces `dplyr` usage
+ [Part 3, Chapter 6 from STAT545](https://stat545.com/dplyr-intro.html)
+ [Chapter 4.3 from another Introduction to Data Science e-book](https://rafalab.github.io/dsbook/tidyverse.html#manipulating-data-frames)
+ [RStudio Cloud exercises for "Work with Data"](https://rstudio.cloud/learn/primers/2)
  + "Working with tibbles" is a general introduction to some useful `tidyverse` and R knowledge
  + "Isolating data with dplyr" and "Deriving information with dplyr" are most relevant to `dplyr`



### The greatest hits of `dplyr` one-table verbs

> According to Dr. Spielman. `dplyr` does *a lot more.*

<br>

+ **Piping syntax** 

  ```{r demo_pipe, eval=FALSE}
  # Without piping
  dplyr_function( data_frame_variable, arguments )
  
  # With piping
  data_frame_variable %>%
    dplyr_function( arguments )
  ```

<br>

+ **[`filter()`](https://dplyr.tidyverse.org/reference/filter.html): Subset _rows_ based on `TRUE`/`FALSE`**

  ```{r demo_filter, exercise=TRUE, exercise.lines = 7}
  # only keep rows where Species == "setosa" is TRUE
  iris %>%
    filter(Species == "setosa")
  
  # only keep rows where the sepal length is greater than 5
  iris %>%
    filter(Sepal.Length > 5)
  ```

<br>


+ **[`mutate()`](https://dplyr.tidyverse.org/reference/mutate.html): Create new _column_**

  ```{r demo_mutate, exercise=TRUE, exercise.lines = 8}
  # create new column called `thiscolumnis5` where the value is 5
  iris %>% 
    mutate(thiscolumnis5 = 5)
  
  # create new column called `sepal_area` as the product of sepal length*width
  iris %>% 
    mutate(sepal_area = Sepal.Length * Sepal.Width)
  ```

<br>


+ **[`distinct()`](https://dplyr.tidyverse.org/reference/distinct.html): Remove duplicate rows**

  ```{r demo_distinct, exercise=TRUE, exercise.lines = 2}
  iris %>% 
    distinct()
  ```

<br>


+ **[`arrange()`](https://dplyr.tidyverse.org/reference/arrange.html): Sort rows according to a given column**

  ```{r demo_arrange, exercise=TRUE, exercise.lines = 6}
  iris %>% 
    arrange(Sepal.Length)
  
  # To arrange in descending order:
  iris %>% 
    arrange(desc(Sepal.Length))
  ```

<br>


+ **[`rename()`](https://dplyr.tidyverse.org/reference/rename.html): Rename columns**

  ```{r demo_rename, exercise=TRUE, exercise.lines = 3}
  iris %>% 
    #       newname     =   oldname    !!!!!!!
    rename(sepal_length = Sepal.Length)
  ```

<br>


+ **[`select()`](https://dplyr.tidyverse.org/reference/select.html): Retain or re-order columns**

  ```{r demo_select, exercise=TRUE, exercise.lines = 8}
  # only keeps those 2 columns
  iris %>% 
    select(Sepal.Length, Sepal.Width)
  
  # remove those two columns and keep the rest
  iris %>% 
    select(-Sepal.Length, -Sepal.Width)
  ```

<br>


+ **[`group_by()`](https://dplyr.tidyverse.org/reference/group_by.html): Create a group for grouped operations**
  + Use `ungroup()` to undo all existing groupings
  
  ```{r demo_group_by, exercise=TRUE, exercise.lines = 10}
  # set up for grouped operations
  iris %>% 
    group_by(Species)
    
  # group and upgroup, as a demonstration:
  iris %>% 
    group_by(Species) %>%
    ungroup()
  ```
  
<br>



+ **[`summarize()`](https://dplyr.tidyverse.org/reference/summarise.html): Create a new data frame that is a _collapsed and summarized_ version**
  
  ```{r demo_summarize, exercise=TRUE, exercise.lines = 10}
  iris %>% 
    summarize(mean_sepal_length = mean(Sepal.Length))
  
  # in combo with group_by!
  iris %>% 
    group_by(Species) %>%
    summarize(mean_sepal_length = mean(Sepal.Length)) %>%
    # to be safe, always ungroup:
    ungroup()
  ```

<br>


+ **[`tally()`](https://dplyr.tidyverse.org/reference/tally.html): Count the number of observations in each group**
  
  ```{r demo_tally, exercise=TRUE, exercise.lines = 5}
  iris %>% 
    group_by(Species) %>% 
    tally() %>%
    # to be safe, always ungroup when you're done
    ungroup()
  ```

<br>


+ **[`count()`](https://dplyr.tidyverse.org/reference/count.html): Handy shortcut to group, tally, ungroup!** 
  
  ```{r demo_count, exercise=TRUE, exercise.lines = 2}
  iris %>% 
    count(Species) 
  ```

