---
title: "Data wrangling round 2"
subtitle: "Data Science for Biologists, Spring 2020"
author: "YOUR NAME GOES HERE"
output: 
  html_document:
    highlight: tango
css: static/sjs-style.css
---

<!-- SETUP CODE CHUNK -->
```{r setup, include=FALSE}
## Change the default figure width and height to your liking, but within reason.
knitr::opts_chunk$set(echo = TRUE, fig.width=6, fig.height=3.5)
library(tidyverse)
library(tidylog) # This must be loaded _AFTER_ tidyverse!
theme_set(theme_classic())  ### If you want to set a global theme, do so here
```


## Instructions

**This assignment must be submitted as an RMarkdown document on Blackboard by _Thursday February 26th at 10:00 am_.** 

#### You will be graded on the following criteria:

+ Did you enter your name appropriately into the YAML front matter?
+ Does the document knit without errors? *Adding `eval=FALSE` to code chunks with errors will not avoid penalty.*
+ Does your code in the R chunks have appropriate *comments*? 
+ Is your submission appropriately _professional_? 
  + *You need to run SPELL CHECK before submitting*, available under `Edit->Check Spelling`.
+ Do all figures have clean axis labels? (Titles are optional). There should NOT be any underscores in axis labels.
+ Did you write your written answers in the <answer>answer style</answer>?
+ Have you addressed the questions as they are asked? Be sure to limit your code and your answers to the scope of the question!
  + If you write code that is not ultimately part of your final answer, you must either: delete it, comment it out, OR move it to a different R chunk with the added chunk options `eval=FALSE, echo=FALSE`. 
+ Is all work your own, including code and written answers?
  + If ANY of your submitted work is not your own (with the exception of what you have been provided), you will receive a 0 on the assignment. You will be reported for a Level 2 or 3 Academic Integrity Violation, depending on the severity of the issue.

<!-- HEY HAVE YOU KNITTED YET? IT'S VERY HARD TO READ INSTRUCTIONS IN RMD! -->


## Questions

For this lab/assignment, you will be using several datasets, all of which are CSV files in the `data/` directory. You'll need to read in the data to use it. **Read in the CSV files for use at the BEGINNING of the question they are used for.** Once a dataset is read in, you don't need to read it in again! Seriously, only read in a dataset once (this counts for grading!!) - otherwise you can accidentally introduce bugs, and bugs are bad. Whenever reading in a dataset, you should ALWAYS IMMEDIATELY PRINT THE DATA (tibbles only print 10 rows by default). This allows readers of your reproducible and sharable document to see the data being used!

```{r}
## read in the DIAMONDS data (assume iris was external and this line reads it in)
diamonds ## PRINT IT OUT!!
```


##### Question 1

bind_rows and bind_cols

##### Question 2

left_join, anti_join, inner_join

##### Question 3
then they'll have to join something to plot it becuase the plotting variables are in two different tibbles.

##### Question 4
missing data (NA)
replace_na
na_if

##### Question 5
try to summarize, realize you can't, think about what to do
drop_na

##### Question 6
if_else() ## MUST BE SAME TYPE


##### Question 7
case_when()


##### Question 8

The `chickwts` dataset contains information on the weight of chicks after being fed different feed supplements. The different feed supplements are labeled casein, horsebean, linseed, meatmeal, soybean, and sunflower in the `feed` column. I have created a new data-frame (`feed_names`), that contains the abbreviated names of different feed supplements. Using one of the `dplyr` join functions, combine the two data-frames so that there is an additional `feed_abbr` column and all of the original columns and rows in chickwts are retained. Which join function is most appropriate to use and why? 

```{r}
head(chickwts)

feed_names <- read.table(text = "
feed feed_abbr
casein cs
whey wh
linseed  ls
meatmeal mm
fishmeal fm
soybean sb
sunflower sf
corn co
wheatbran wb
", head = T)

# R code goes here
```

HINT: Here are some functions you'll probably want to use when planning your approach to this question:
+ `rename` to change column names
+ `case_when` to create recoded columns (Hint2: creating new columns? sounds like this might necessitate another function as well! what could it be?)
+ `bind_rows` to combine two tibbles with the same column names (and data types in those columns!)

woops somebody else screwed up their data 
they put into two tibbles! your task is to combine them.
second tibble has feeds by their abbr, so you need to use case_when to mutate these.
also weight should be called chick_weight so they'll need to use rename

##### Question 9

oh god they really screwed up. now they have a second set of information about nutritional content for the feeds. 


##### Question 10
Create a visualization that shows the relationship between **median weight** for each feed and the **nutritional content** of each feed type as measured in something totally different. To achieve a dataset you can visualize you'll have to wrangle!







