---
title: "Data wrangling round 2"
subtitle: "Data Science for Biologists, Spring 2020"
author: "YOUR NAME GOES HERE"
output: 
  html_document:
    highlight: tango
css: static/sjs-style.css
---

<!-- SETUP CODE CHUNK -->
```{r setup, include=FALSE}
## Change the default figure width and height to your liking, but within reason.
knitr::opts_chunk$set(echo = TRUE, fig.width=6, fig.height=3.5)
library(tidyverse)
library(tidylog) # This must be loaded _AFTER_ tidyverse!
theme_set(theme_classic())  ### If you want to set a global theme, do so here
```


## Instructions

**This assignment must be submitted as an RMarkdown document on Blackboard by _Thursday February 26th at 10:00 am_.** 

#### You will be graded on the following criteria:

+ Did you enter your name appropriately into the YAML front matter?
+ Does the document knit without errors? *Adding `eval=FALSE` to code chunks with errors will not avoid penalty.*
+ Does your code in the R chunks have appropriate *comments*? 
+ Is your submission appropriately _professional_? 
  + *You need to run SPELL CHECK before submitting*, available under `Edit->Check Spelling`.
+ Do all figures have clean axis labels? (Titles are optional). There should NOT be any underscores in axis labels.
+ Did you write your written answers in the <answer>answer style</answer>?
+ Have you addressed the questions as they are asked? Be sure to limit your code and your answers to the scope of the question!
  + If you write code that is not ultimately part of your final answer, you must either: delete it, comment it out, OR move it to a different R chunk with the added chunk options `eval=FALSE, echo=FALSE`. 
+ Is all work your own, including code and written answers?
  + If ANY of your submitted work is not your own (with the exception of what you have been provided), you will receive a 0 on the assignment. You will be reported for a Level 2 or 3 Academic Integrity Violation, depending on the severity of the issue.

<!-- HEY HAVE YOU KNITTED YET? IT'S VERY HARD TO READ INSTRUCTIONS IN RMD! -->


## Questions

Perform all exercises using the `%>%` operator. For this assignment, you will be working with TWO datasets which can be found in the `data/` directory in your RStudio Cloud Project:
 
+ `annual_AIDS_death.csv`

from claus, meh


**This homework is due on Feb. 19, 2019 at 4:00pm. Please submit as a PDF file on Canvas.**

**Problem 1: (4 pts)** The following two data tables contain information about how many male and female passengers traveling on the Titanic in different classes survived or died. The data-frame `survived` contains information about passengers that survived, and the data-frame `deceased` contains information about passendgers that died. Using the `dplyr` and `tidyr` packages, make these data-frames tidy and then combine them into a single data-frame. Make sure that your final data-frame has a `survival_status` column indicating which data-frame the observations originally came from. HINT: You can use the `bind_rows` function to add rows from one data-frame onto another as long as both data-frames have identical column names.

```{r}
survived <- read.table(text = "
class  male female
1st    57    140
2nd    14     80
3rd    75     76
crew  192     20
", head = T)

deceased <- read.table(text = "
class  male female
1st   118      4
2nd   154     13
3rd   387     89
crew  670      3
", head = T)

# R code goes here
```

Using the data-frame you created above, compute the total number of passengers that survived and that did not survive.

```{r}
# R code goes here
```

**Problem 2: (3 pts)** The `chickwts` dataset contains information on the weight of chicks after being fed different feed supplements. The different feed supplements are labeled casein, horsebean, linseed, meatmeal, soybean, and sunflower in the `feed` column. I have created a new data-frame (`feed_names`), that contains the abbreviated names of different feed supplements. Using one of the `dplyr` join functions, combine the two data-frames so that there is an additional `feed_abbr` column and all of the original columns and rows in chickwts are retained. Which join function is most appropriate to use and why? 

```{r}
head(chickwts)

feed_names <- read.table(text = "
feed feed_abbr
casein cs
whey wh
linseed  ls
meatmeal mm
fishmeal fm
soybean sb
sunflower sf
corn co
wheatbran wb
", head = T)

# R code goes here
```

  


