---
title: "ANSWERS: Working with untidy and tidy data"
subtitle: "Data Science for Biologists, Spring 2020"
author: "YOUR NAME GOES HERE"
output: 
  html_document:
    highlight: tango
css: static/sjs-style.css
params:
  data_path: "data/"
---

<!-- SETUP CODE CHUNK -->
```{r setup, include=FALSE}
## Change the default figure width and height to your liking, but within reason.
knitr::opts_chunk$set(echo = TRUE, fig.width=6, fig.height=3.5)
library(tidyverse)
library(kableExtra) # helpful to print fancy tables in RMarkdown
theme_set(theme_classic())  ### If you want to set a global theme, do so here
```


# <mark>Questions</mark>

#### Preface Question

Look at the YAML front matter: there is something NEW! Read about params [here](https://rmarkdown.rstudio.com/lesson-6.html). In this case we set a global parameter for the RMarkdown file that corresponds to the path where data lives! We can proceed to use this as a variable like so (always `params$`, like a column in a data frame!). Use this variable for your paths, e.g. `read_csv(paste0(params$data_path, <NAMEOFFILE>))`
```{r}
params$data_path
```

#### Question 1

```{r}
# Code chunk to read in and examine the data with head()
```


```{r}
# Code chunk to convert data to tidy
```
#### Question 2


```{r}
# Code chunk to plot the data
```

<answer>
Replace this text with your answer.
</answer>

#### Question 3


```{r}
# Code chunk to read in and examine the data with head()
```


```{r}
# Code chunk to convert data to tidy and combine with question 1 data
```

```{r}
# Code chunk to plot the data
```

<answer>
Replace this text with your answer.
</answer>

#### Question 4


```{r}
# Code chunk to read in and examine the data with head()
```


```{r}
# Code chunk for distributions (violin, sina, etc.)
# If sina make sure to load `ggforce()` in SETUP CHUNK!
```

```{r}
# Code chunk for data un-tidying and scatterplot
```

<answer>
Replace this text with your answer.
</answer>

#### Question 5


```{r}
# Code chunk to read in and examine the data with head()
```


```{r}
# Code chunk for tidying
```


#### Question 6

```{r}
# Code chunk for allllll the wrangling and tidying
# should END UP as a single piped command - but build it up LINE AT A TIME!
```


#### Question 7


```{r}
## defines each column as an array
tibble(
  color        = c("blue", "green", "turquoise"),
  male         = c(25, 36, 52),
  female       = c(28, 17, 33),
  species_abbr = c("sp1", "sp1", "sp1")
) -> frogs_sp1

## defines tibble in a format like the tibble would end up appearing
## spacing is helpful!
tribble(
  ~color, ~male, ~female, ~species_abbr,
  "orange", 13,   17,    "sp2",
  "red",    45,   36,    "sp2",
  "pink",  31,    25,    "sp2"
) -> frogs_sp2

## Like strategy one
tibble(
  species_name = c("Frogus firstus", "Frofus secondus"),
  species_abbr = c("sp1", "sp2")
) -> species_names

### ADD CODE HERE TO PRINT OUT THESE THREE DATASETS!!! YOU HAVE TO LOOK AT THEM TO USE THEM!!
```



```{r}
# Code chunk to tidy and combine the three datasets
```

```{r}
# Code chunk to for dplyr task 1
```

```{r}
# Code chunk to for dplyr task 2
```

```{r}
# Code chunk to for dplyr task 3
```



#### Question 8

```{r}
# Code chunk for kable()
```

#### BONUS QUESTION

*Add your own chunks if you choose to do this adventure.*

<answer>
Replace this text with your answer.
</answer>