---
title: "Tidy Tuesday Week 2"
subtitle: "Measles"
author: "Shemaiah S. Sotrah"
date: "February 30, 2020"
output: 
  html_document:
    highlight: tango
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(maps)
```

```{r, message=FALSE}
measles <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-25/measles.csv') #read in measles
measles #view measles data frame
```

```{r, message=FALSE}
measles %>%
  group_by(state) %>%
  summarise(mean_mmr = mean(mmr)) %>% #mean mmr for each state
  rename( region = state)-> measles2 #state becomes region to join datasets

map_data("state") -> usa_map #second dataset

measles2 %>%
mutate(region = str_to_lower(region)) %>% #need to all be lower to join data
left_join( usa_map, measles2, by = "region") -> final_map 
 
ggplot(final_map, aes( x = long, y = lat, group = group, fill = mean_mmr)) + 
  geom_polygon (color = "black") +
  coord_fixed(1.3) + #maintains ratio of map by x and y
  scale_fill_gradientn(colours = rev(rainbow(9)),breaks = c(0, 25, 50, 75, 100)) + #continuous and specific break, otherwise cut off
  theme_classic() +
  labs(fill = "Mean MMR",
       x = "Longitude",
       y = "Latitude",
       title = "School's Measles, Mumps, and Rubella Vaccination Rate in the USA")

ggsave("mmr_usa.jpeg")
                
```                

  This plot actually took me a LONG time. It was hard to find blog posts where tidyverse was used instead of base R. However, it was good because reading many different posts helped me to better learn the "programming lingo. Overall, even though this may not be a great plot or even perfectly what I had in mind, it was rewarding when it was finished. I think it is so cool that you can put data on a map. I remember seeing plots like this and thinking that whoever made them must be a genius. Now, I can do it!
           