---
title: 'Instructions: Homework #8'
author: "Data Science for Biologists, Fall 2021"
date: "Complete the template Rmd and submit to Canvas on Wednesday 11/3/21 by 11:59 PM"
output: 
  html_document:
    theme: lumen
    highlight: tango
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, include=F, fig.width = 6, fig.height = 4, message = F, warning = F)

library(tidyverse)
coffee_ratings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-07/coffee_ratings.csv')
coffee_ratings <- coffee_ratings %>%
  filter(total_cup_points != 0)
```


```{css, echo=F}
ul :last-child, ol :last-child,ul :first-child, ol :first-child {
    margin-bottom: 0;
    padding-bottom:0;
}

blockquote {
  margin-left: 50px;
  border-left: 5px solid #DF4469;
  font-size:1em;
}
```


### Obtaining and setting up the homework

+ Obtain the homework template from your RStudio Cloud class project by running the following code in the R Console:

  ```{r, eval=FALSE, echo=T, include=T}
  library(ds4b.materials) # Load the class library
  launch_homework(8)      # Launch Homework 8
  ```

+ You must set an _RMarkdown theme and code syntax highlighting scheme_ of your choosing in the YAML front matter. These links will help you:
  + Choose your favorite _theme_ among the **pre-packaged themes** (ignore everything below "Even More Themes") shown at [this link](https://www.datadreaming.org/post/r-markdown-theme-gallery/)
  + Choose your favorite _syntax highlighting_ among these options at [this link](https://www.garrickadenbuie.com/blog/pandoc-syntax-highlighting-examples/)
+ **Make sure your Rmd knits without errors before submitting.** If it does not produce an HTML output, this means it does not knit. _DO NOT SKIP THIS STEP!_ Ensuring code runs without errors is MORE IMPORTANT than writing code in the first place.
  + If there are errors in your code, you should *comment out* the code so that it does not actually run. This is **BETTER** than keeping the buggy code in there without commenting out - _it shows me you attempted the code, but understood that it didn't work properly._ Partial credit will come to you! But, if you leave buggy code in, then the Rmd will not knit and there will be deductions.
+ As always, you are encouraged to work together and use the class Slack to help each other out, but you must submit YOUR OWN CODE.
<br><br>


<br><br>

# The actual instructions


## Part 1: Reading in the data

In the top chunk named `read_data`, first... 

+ Read in the `coffee_ratings` dataset *directly from the URL where it lives on the internet* (not from the file). _This is done for you._
+ Then, since we know there is an error in this dataset where 1 row has a point score of 0, let's remove it before proceeding. Filter the data to only keep rows rows where `total_cup_points != 0`, and save the data to `clean_coffee_ratings`.**You should use this tibble, `clean_coffee_ratings`, for the whole homework.** Feel free to change the variable name if you like as long as it remains informative and _not the same as_ the raw data `coffee_ratings`.

## Part 2: Go for it!!

You will be conducting exploratory data analysis on the coffee ratings dataset by answering **five questions** about the coffee data using a combination of wrangling and visualization. Three of these questions are asked for you, and you ask the other two questions. Each question needs at least _some wrangling_ with one or more `dplyr` and/or `tidyr` functions, its own plot (although question 3 has two plots!), and a brief answer in 1-3 sentences. Please follow the given template to organize your code (as you did in Homework #5). You can style your plots _however you want_ as long as you ensure professional labeling (no underscores!) of all axes and legend titles. 

**Importantly, for every written answer you give there MUST BE corresponding code.** For the first three questions templated for you, you must conduct calculations as described.

**Finally, YOU MUST SPELL CHECK. Seriously.**

<br>

#### Specific tasks you have to do as part of questions 1-3

**Question 1**

+ Your code should make sure to accomplish the following goals:
  + Use wrangling to reveal a tibble of the _standard deviation_ of each processing method's uniformity. This tibble should have two columns named `processing_method` and `sd_uniformity`.
  + Create a figure to visualize the spread of the uniformity distributions across non-`NA` processing methods
+ **Tips**
  + Your wrangling is totally separate from your plotting. Wrangling needs *three* `dplyr` verbs.
  + _Do not make a bar plot!_ You should plot the *distributions* (overlapping density, faceted histogram, boxplot, jitter, violin, etc). of uniformity across processing methods. The idea here is that the wrangling and the plot are *complementary ways to think about data spread.* The wrangling tells us precisely what the spread is, and the plot allows us to derive a visual sense of how spread out of the data is.

<br>

**Question 2**

+ Your code should make sure to accomplish the following goals:
  + Reveal a tibble of how many Robusta coffees are grown in countries that grow Robusta coffees (i.e. no rows with "0"). This tibble should have two columns named `country_of_origin` and `number_robusta` 
  + Create a figure to visualize these counts. 
    + For some bonus points, make a [waffle plot](https://liamgilbey.github.io/ggwaffle/)! If you want to pursue this bonus option, you first need to install the package `ggwaffle` with this code (copy/paste into Console): `remotes::install_github("liamgilbey/ggwaffle")`.
+ **Tips**
  + Your wrangling needs *two* `dplyr` verbs. 
  + You can pipe the wrangling into the plot if you want, but not strictly required.
  
<br>

**Question 3**

+ Your code should make sure to accomplish the following goals:
  + Use wrangling to reveal a tibble of the _mean_ of each color's moisture, excluding unknown (`NA`) colors. This tibble should have two columns named `color` and `mean_moisture`.
  + Create TWO FIGURES for this question: The first should show the moisture distributions across non-`NA` colors, and the second should show the the literal mean values of each moisture distribution (hint: barplot!). You can make both figures in the same chunk, or create an additional (named!) chunk as needed.
+ _Your answer_ should specifically address which figure, in your opinion, is most helpful for answering the question _and why._
+ **Tips**
  + Your wrangling needs *three* `dplyr` verbs. 
  + The first plot you have to make here is a bar plot (using `geom_col()`! Think about why this geom!).
  + The second plot you have to make here is literally only asking you to plot `moisture` across (non-`NA`) categories of `color`. It has nothing to do with the means that were necessary to calculate for making the first plot.
  + The purpose of this question is to see the difference between looking at numeric distributions as "just the mean of all those numbers" vs "the full distributions of the numbers themselves." So, your answer should discuss which is more informative for understanding GENERALLY how moisture differs across color categories.
  + I highly recommend _not making a boxplot_ for your second plot. It might be a neat adventure to take along the way though so you can see why I recommend no boxplots here (it will also teach you something about pitfalls of boxplots!).
