---
title: 'Template: Homework #8'
author: "Data Science for Biologists, Fall 2021"
date: "Submit this completed Rmd file to Canvas on Wednesday 11/3/21 by 11:59 PM"
output: 
  html_document:
    theme: lumen
    highlight: tango
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 6, fig.height = 4)

# Always load libraries in the setup chunk:
library(tidyverse) # allows you to use ggplot2, readr, forcats, etc. 
```


```{r read_data, include=FALSE}
# Read the coffee data from URL
coffee_ratings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-07/coffee_ratings.csv')

# Your task: Remove the incorrect row from the dataset and save to a new tibble to use in the homework

```

<br><br>

#### Question 1: How does the spread of coffee `uniformity` distributions compare among different coffee processing methods, excluding unknown (`NA`) processing methods? 

Replace this text with your answer to question 1.

```{r question1}

```

<br>


#### Question 2: Which countries grow Robusta coffees, and how many Robusta coffees are grown in each of these countries?

Replace this text with your answer to question 2.

```{r question2}

```

<br>

#### Question 3: Does there appear to be an association between `color` and `moisture` across coffees? In other words, do certain colors have a higher or lower moisture on average?

Replace this text with your answer to question 3.

```{r question3}

```


<br>

#### Question 4: Replace this text with your question. 

Replace this text with your answer to question 4.

```{r question4}

```



<br>

#### Question 5: Replace this text with your question. 

Replace this text with your answer to question 5.

```{r question5}

```

