---
title: 'Template: Homework #5'
author: "Data Science for Biologists, Fall 2021"
date: "Submit this completed Rmd file to Canvas on Wednesday 10/13/21 by 2 PM"
output: 
  html_document:
    theme: lumen
    highlight: tango
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 6, fig.height = 4)

# Always load libraries in the setup chunk:
library(tidyverse) # allows you to use ggplot2, forcats, etc. 
library(ds4b.materials) # allows access to all datasets dataset
```


## Example dataset: `iris`

#### What is the relationship between sepal lengths and sepal widths, for each specices of iris?

All species show a positive relationship between sepal length and sepal width. Because confidence intervals for `virginica` and `versicolor` are overlapping, it is likely they have the same relationship between sepal lengths and widths. However, the `setosa` relationship has a steeper slope.

```{r iris_plot1}
ggplot(iris) + 
  aes(x = Sepal.Length, 
      y = Sepal.Width, 
      # color aes allows to see relationships for each species separately
      color = Species) + 
  geom_point() + 
  # add trendline to see relationship more clearly
  geom_smooth(method = "lm") + 
  # use a non-default color palette
  scale_color_brewer(palette = "Set1") +
  # add professional labels
  labs(x = "Sepal Length", y = "Sepal Width") + 
  # set a non-default theme
  theme_minimal()
```

<br><br>


## Dataset `replace this text with the chosen dataset, keeping the backticks`

#### Replace this text with your first question, ensuring it remains a proper Level 4 header in the knitted HTML.

Replace this text with your answer to the first question.

```{r}

```

<br>

#### Replace this text with your second question, ensuring it remains a proper Level 4 header in the knitted HTML.

Replace this text with your answer to the second question.

```{r}

```

<br><br>


## Dataset `replace this text with the chosen dataset, keeping the backticks`

#### Replace this text with your first question, ensuring it remains a proper Level 4 header in the knitted HTML.

Replace this text with your answer to the first question.

```{r}

```

<br>

#### Replace this text with your second question, ensuring it remains a proper Level 4 header in the knitted HTML.

Replace this text with your answer to the second question.

```{r}

```

<br><br>


## Dataset `replace this text with the chosen dataset, keeping the backticks`

#### Replace this text with your first question, ensuring it remains a proper Level 4 header in the knitted HTML.

Replace this text with your answer to the first question.

```{r}

```

<br>

#### Replace this text with your second question, ensuring it remains a proper Level 4 header in the knitted HTML.

Replace this text with your answer to the second question.

```{r}

```

<br><br>