---
title: "EXAMPLE PROPOSAL"
subtitle: "Data Science for Biologists, Fall 2020"
author: "Stephanie J. Spielman"
output: 
  html_document:
    theme: simplex
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo = TRUE, include = TRUE)
# Load all libraries you use HERE
library(tidyverse)
```

<br>

### Description

For my project, I will be analyzing a [dataset of coffee ratings](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-07-07/readme.md) from the `#tidytuesday` repository. This dataset contains information about coffees and their ratings from around the world. I chose this dataset because I love (too much?) to drink coffee!

<br>

### Data

```{r read-data}
coffee_ratings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-07/coffee_ratings.csv')

coffee_ratings
```


<br>

### Modeling question

For my model question, I will use a linear regression to ask extent a coffee's processing method can explain its flavor rating. The predictor variable is `processing_method`, and the response variable is `flavor`. 

<br>

### Practice plot

```{r practice-plot, fig.width = 6, fig.height = 4}
ggplot(coffee_ratings, aes(x = species, y = aroma, fill = species)) + 
  geom_violin(alpha = 0.8) + 
  scale_fill_manual(values=c("purple", "chocolate4")) + 
  theme_bw()
```

### Github account

My github account is named `sjspielman`.

