---
title: 'Homework #2'
author: "Data Science for Biologists, Spring 2021"
date: "Due as an R Script to Canvas on Tuesday 9/14/21 by 1 PM"
output: 
  html_document:
    theme: lumen
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Instructions

For this homework, you will be writing your very first R script and learning how to engage with RStudio and the `ds4b.materials` R library from Dr. Spielman. The goal of this homework is to provide you with a concrete opportunity to gain initial comfort with the fundamental tools used in this class and enable your future success for the rest of the semester.

The literal step-by-step instructions that you should take to complete this homework are in these videos. Please watch them in this order, and follow instructions as you go:

[Link to video(s) will go here]


### What NOT to do vs what TO do

> The following code chunks provide examples of how *NOT TO* and how *TO* answer a hypothetical homework question using code. Notice how there are many ways to correctly achieve the right result! Welcome to coding :)

**Hypothetical homework question: What is the length of an array with values 4, 6, and 10?**

<br>

**NO!** This code is not acceptable, largely because it is not code - it is just the number 3. There is no code used at all to answer the question. Even if it is "easy" to do something in your head, you must use code to answer it. *It also will not produce any output from a script since it is not formally printed.*
```{r eval=FALSE}
3
```

<br><br>
**NO!** This code is not acceptable, because even though code was used at some point, the code itself does not directly produce the output. This example represents a situation where you might run `length(c(4,6,10))` in the R Console, or run `length(c(4,6,10))` in the chunk but later delete or comment it out. You might look at the output and say, "the code calculated 3, so the answer is 3." But, this is not ok, since the code you submit itself must produce the number 3, and we are missing the print statement.

```{r eval=FALSE}
#length(c(4,6,10))
3
```

<br><br>
**NO!** This is getting much closer, since we now have the fully code needed to generate the output. However, nothing is actually printed, so the output is not revealed. It is critical to always reveal the output!
```{r eval=FALSE}
answer <- length(c(4,6,10))
```


<br><br>
**YES!** This "fixes" the previous "NO!" example by revealing the output.
```{r}
answer <- length(c(4,6,10))
print( answer )
```
<br><br>
**YES!** Only code is used to produce the output.
```{r}
# YES!
print( length(c(4,6,10)) )
```

<br><br>
**YES!** Only code is used to produce the output. This strategy uses improved coding style by defining a _variable_ for the array, and then using `length()` to calculate the length of the variable.
```{r}
# YES!
my_array <- c(4,6,10)
print( length(my_array) )
```
<br><br>
**YES!** Only code is used to produce the output. This strategy uses improved coding style by defining a _variable_ for the array, and then defining another variable to contain the output from the `length()` calculation.
```{r}
# YES!
my_array <- c(4,6,10)
the_length <- length(my_array)
print( the_length )
```
<br>
**Unless the instructions explicitly ask you to define certain variables, all "YES!" code versions here are equivalent and entirely accceptable!**

<br><br>

### The literal homework questions:

Again, all answers must have corresponding code, and answers must be directly derived from code. Your code must include comments. Use the example code as provided in the videos to learn how you are expected to comment!

<br>


**Example Question discussed answered in videos**: Define a variable `w` as the value 10, and print the variable.

<br>

1. Define a _character_ variable named `sport` as a string value of your favorite sport (or other activity if you hate sports!). Print the variable using the `print()` function.

2. Define the variable `a` with the numeric value 36. Use the `sqrt()` function to calculate the square root of the `a` variable, and print the result of this operation using the `print()` function. 

3. Define an array variable called `mammal_vectors` to contain these strings: "camel", "civet", "mink", "bat", "ferret", and "pangolin" (these are all mammals known to be vectors for either SARS-CoV-2 or one of its close relatives SARS and MERS). Use the function `length()` to determine how long the array is, and print the result using the `print()` function. 

4. Create an array variable called `some_numbers` that contains 4 unique numbers of _your choosing_. Use the function `log` to calculate the natural log of all values in the array. This operation will produce a new array containing the natural log of the numbers you provided; print this new array using the `print()` function.
+ Note that if you provide negative numbers or 0, you may get answers like `NnN` ("not a number") or `Inf` (infinity). That's completely fine! It's the correct result of the calculation.

5. The templated code for this question has a bug! The tasks the code is meant to accomplish are provided in comments in the homework template. Fix the bugs in the templated code according to instructions in the homework videos.

6. Again, the templated code for this question has a bug! The tasks the code is meant to accomplish are provided in comments in the homework template. Fix the bugs in the templated code according to instructions in the homework videos.