---
title: "Demonstration Rmarkdown document"
author: "Stephanie J. Spielman"
date: "3/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Header level 1

## Header level 2

##But if you forget to add a SPACE after the hashtag(s), your header won't work :(

### Header level 3

#### Header level 4

##### Header level 5

<!-- Hashtags are NOT COMMENTS IN MARKDOWN!! They are only comments in R!!!. To get a comment here, you need to use HTML comments. You'll notice this sentence does NOT APPEAR in the knitted document. -->


We can style text with **bold**, *italics*, and `code style`. Surrounding text with dollar signs gives "math mode," like this: $Y=X$. Look at this fanciness: $R^2 = 0.89$. We can include [links](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-07-07/readme.md) pretty easily, too.

> Blockquotes are also pretty cool.

+ bullet list
+ more bullet
  + a nested bullet!
    + more nesting! These nested bullets are TABBED in. 

1. Or, have a NUMBERED list.
2. Yup, it works!
  a. But gets funky when you try nesting. It's often easier to use *bullets* if you want to nest a list.
  
For starting a new LIST (bullets or numbers), you need to have a NEW LINE before it. Otherwise, it's a mess!
+ this bullet doesn't work.

+ now the bullet works since there's a blank line above it!


See? I bet you think this text
is all written on the same line, 
but it's not. Markdown is not Word. It doesn't care where you type "enter."
Do you REALLY want to enter? Use `<br>` which is actual HTML. All actual HTML will be converted to..it's actual HTML project.
<br>
Ahhhaaaa now we;re on the next line. Want more spacing lines? More `<br>`!
<br><br>
Ah, a nice big break.


## markdown and Rmarkdown resources

+ [markdown cheatsheet](https://github.com/adam-p/markdown-here/wiki/Markdown-Cheatsheet)
+ [another markdown cheatsheet](https://commonmark.org/help/)
+ [Rmarkdown cheatsheet](https://rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf)
+ [Rmarkdown tutorial](https://ourcodingclub.github.io/tutorials/rmarkdown/#what)
+ [Rmarkdown guide](https://www.dataquest.io/blog/r-markdown-guide-cheatsheet/)
+ [R4DS Rmarkdown chapter](https://r4ds.had.co.nz/r-markdown.html#r-markdown)
+ [Official RMarkdown Reference](https://rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf)
+ [Literally everything there is to know about Rmarkdown, and then some. Really overwhelming, but really comprehensive.](https://bookdown.org/yihui/rmarkdown/)
+ [Yihui Xie, author of Rmd, is a hilarious genius.](https://rstudio.com/resources/rstudioconf-2020/one-r-markdown-document-fourteen-demos/)


## Introducing R "code chunks"

This code chunk defines `x` as the value 5 and prints out the value.
```{r}
# define x as 5
x <- 5
# print the value of x
x
```

This code chunk, which appears *AFTER* the code chunk that defined `x<-5`, will print out 5! Code chunks will be run sequentially (in order from the top down in the document) when you **KNIT.**
```{r}
x
```

**But, we really should name our chunks.** Naming chunks makes it much easier to navigate your document - imagine having 50 chunks in one Rmarkdown document and having to find the right one you need to work with! **NAME THEM, NAME THEM!!**
```{r name_of_chunk}
y <- 10
z <- x + y
z
```

You'll notice above that the *output* from the code appears separately. Don't like that look? You can change the *setting* `collapse`, whose default is `FALSE`.
```{r demo_collapse_true, collapse=TRUE}
print(z)
```

You'll see in the _Rmd file itself_ (not in the knitted output) that the code chunk setting must be organized like this: `{language name, settings, more settings, more settings}`: The first item is which language the chunk is in (for us, always `r`!!), and the second item is the chunk's name, and there is NO COMMA between the language and name. Then, we use commas to separate any additional settings such as `collapse=TRUE`.

If you want to apply any setting like `collapse=TRUE` automatically to ALL code chunks (rather than setting each code chunk separately), add the setting into the OPENING `setup` chunk specifically inside this line:
`knitr::opts_chunk$set(SET OPTIONS TO TRUE IN HERE THAT YOU WANT TO BE APPLIED TO EVERYTHING)`. Settings that are commonly set either in `setup` or for individual chunks include the following, **which you can also set up interactively with the chunk's gear icon!**

+ `include`
  + `FALSE`: The *code* in the code chunk *will not* be displayed in the knitted HTML
  + `TRUE`: The *code* in the code chunk *WILL* be displayed in the knitted HTML
+ `echo`
  + `FALSE`: The *output* from the code chunk *will not* be displayed in the knitted HTML
  + `TRUE`: The *output* from the code chunk *WILL* be displayed in the knitted HTML
+ `collapse`
  + `FALSE`: The *output* from the code chunk will be displayed separately from the code in the knitted HTML
  + `TRUE`:  The *output* from the code chunk will be displayed as embedded comments with the code in the knitted HTML
+ `warning` and `message`
  + `FALSE`: warnings or messages (depending which set) resulting from the code *will be suppressed* in the knitted HTML
    + For example, remember the annoying message from `geom_histogram()` about bins?
  + `TRUE`: warnings or messages (depending which set) resulting from the code *will be included* in the knitted HTML
+ **Very important:** `fig.width` and `fig.height`: The default width and height (in) for figures displayed in the knitted HTML. **It is common to need to customize this for EACH code chunk.** Setting an overall default is helpful, but all figures are different and may need their own sizes!

## Get fancier!!

You can choose the *Rmarkdown theme* and *code highlight style* by modifying the YAML.


#### Modifying the document theme

Change the `output: html_document` section of the YAML to look like this (notice the _tabbing_ - without tabbing, it won't work!)

Example:
````markdown
output: 
  html_document:
    theme: cerulean
````

There are [lots of awesome themes](https://www.datadreaming.org/post/r-markdown-theme-gallery/)!


#### Modifying the document theme

Change the `output: html_document` section of the YAML to look like this: 

Example:
````markdown
output: 
  html_document:
    highlight: pygments
````

There are [lots of awesome highlight styles](https://www.garrickadenbuie.com/blog/pandoc-syntax-highlighting-examples/#:~:text=The%20R%20Markdown%20book%20lists,null%20to%20prevent%20syntax%20highlighting.)!



#### How about a table of contents?

Just like for `theme` and `highlight`, reformat the YAML to have `toc: true` under `html_document`.

````markdown
output: 
  html_document:
    toc: true
````
For a *floating* table of contents, set both `toc` and `toc_float` to `true`:

````markdown
output: 
  html_document:
    toc: true
    toc_float: true
````

#### Yes of course, you can have it all

````markdown
output: 
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float: true
````


### Helpful hints

+ To compile the final product, you must click the "Knit" button. 
  + R chunks will be run "brand new" in the order you wrote them
+ You can run single code chunks with the green button in their top right corner. When doing this...
  + Code chunks will NOT communicate. If you click "Run" for chunk #3, R will NOT AUTOMATICALLY RUN the previous 2 code chunks.
  + This is just like pressing "Run" from a script: Copy/paste into the Console.
  + Single chunks running successfully is NOT THE SAME as your document knitting successfully!!


### Prefer a PDF output to HTML?

+ Option 1 (*worse option*): Replace `html_document` with `pdf_document` in the YAML, ie: `output: pdf_document`
  + This will give a somewhat differently formatted document! It will look different from the HTML version.
  + Not all the highlights/themes will port to PDF this way.
+ Option 2 (*better option*): Run the code in Console: `pagedown::chrome_print("name_of_rmarkdown_file.Rmd")`
  + This will give a PDF output that looks as close as possible to the HTML outputted version.
  + Requires that you have installed the `pagedown` package.
  