---
title: "Using themes in `ggplot2`"
date: "Stephanie J. Spielman"
author: "Data Science for Biologists, Fall 2020"
output: 
  html_document:
    theme: united
    toc: true
    toc_float: true
    toc_levels: 4
    css: ../static/sjs-style.css
editor_options: 
  chunk_output_type: console
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  fig.width = 4,
  fig.height = 3
)
library(ggplot2)
```
<br>

> WORD TO THE WISE: You will often be tempted to google how to use `ggplot2`. This is absolutely the right thing to do, but be careful!! Not all internet posts are created equally, and _many webpages or blogs may use DIFFERENT VERSIONS of `ggplot2`_. When asking the internet for help, always make sure you can locally (on your computer/in RStudio Cloud) run and understand little code tidbits you find. Remember: just because a page looks helpful doesn't mean it is.

We use the term *theme* to refer to all non-data components of a plot, including things like "where is the title placed?", "how large is the axis text?", "does the plot have a grid and how spaced out is the grid?", "is there a background fill to the plot canvas itself?", etc.

Themes are comprised of several different types of elements:

+ **`element_text()`: How do certain text elements appear?**
  + For example, `axis.title` (axes titles) is an `element_text()`. 
  + Among other options, you can change their:
    + `color`
    + `size`
    + `face` : "plain", "italic", "bold", or "bold.italic)
    + `angle`
    + `hjust` : "horizontal justification" where 0 is far left, 0.5 is middle, and 1 is far right, and anything in between!
    + `vjust` : "vertical justification" where 0 is bottom, 0.5 is middle, and 1 is top, and anything in between!)
    
<br>  
    
+ **`element_line()`: How do certain line elements appear?**
  + For example, `axis.line` (axis lines themselves) is an `element_line()`. 
  + Among other options, you can change their `color`, `size`, `linetype` ("solid", "dashed", "dotted", "dotdash", "longdash", and "twodash"), and `lineend` ("round", "butt", "square").

<br>  
  
+ **`element_rect()`: How do certain _borders and background_ elements in the plot appear?** 
  + For example, `plot.background` (the background/"canvas" appearance of the plot) is an `element_rect()`. 
  + Among other options, you can change their:
    + `color`
    + `fill`
    + `size` (of the outline)
    + `linetype` ("solid", "dashed", "dotted", "dotdash", "longdash", and "twodash")
  
<br>

All theme components can be fully customized, as is overwhelmingly documented [here](https://ggplot2.tidyverse.org/reference/theme.html). As we have previously seen, there are also several [existing complete themes](https://ggplot2.tidyverse.org/reference/ggtheme.html) in `ggplot2`. Now we'd like to learn how to further customize our plots with our own themes.


All examples below will modify a version of this plot which uses the default `theme_gray()` with no further customization:

```{r}
ggplot(iris, aes(x = Sepal.Length)) + 
  geom_histogram()
```

## Specifying a complete theme

For a given plot, simply add the theme of interest. For example:

```{r}
ggplot(iris, aes(x = Sepal.Length)) + 
  geom_histogram() + 
  theme_classic() # Set the theme as theme_classic() by adding
```





## Customizing a theme

When customizing a theme, you need to:

+ Search through [the documentation](https://ggplot2.tidyverse.org/reference/element.html) (literally) to find how `ggplot2` refers to the particular plot component you want to change. 
+ Use [the documentation](https://ggplot2.tidyverse.org/reference/element.html) (sensing a pattern?) to determine whether that component goes with `element_text()`, `element_line()`, or `element_rect()`. 
  + Or, decide if you want to *entirely remove* that particular plot component. Stay tuned...
+ *Add it* to your plot inside the `theme()` function.
  + If you are also specifying a complete theme in your plot code but want to change certain components, make sure to add these components *after* you specify the complete theme. Otherwise, the complete theme settings will *override* your customizations.
+ There is a *very very VERY* useful function you can use as part of the code call `rel()` which will help you to change sizes of different elements *relative to the baseline*. `rel(1.1)` means 10% larger than default; `rel(0.9)` means 90% the size of the default. 

### Quick examples of `element_text()`

Let's see some examples using `axis.text`^[As you will notice when perusing the documentation, there is also `axis.text.x` and `axis.text.y` which can be used to separately customize X and Y axis text.], which controls both X and Y axis *text* (aka the labels along the axes, NOT the title). Customizing `axis.text` must be paired with `element_text()`, as it is text!

<br><br>

Make the axis text 50% larger than its default:
```{r}
ggplot(iris, aes(x = Sepal.Length)) + 
  geom_histogram() + 
  #theme( plot-thing-to-change = element_<YPE>(things to customize about it) )
  theme(axis.text = element_text(size = rel(1.5)))
```

<br>
Make the axis text 50% smaller than its default:
```{r}
ggplot(iris, aes(x = Sepal.Length)) + 
  geom_histogram() + 
  theme(axis.text = element_text(size = rel(0.5)))
```


<br>
Make the axis text red, *and* 15% larger than default:
```{r}
ggplot(iris, aes(x = Sepal.Length)) + 
  geom_histogram() + 
  theme(axis.text = element_text(color = "red", size = rel(1.15)))
```

<br>
Let's change a whole bunch of things! Change baseline theme to `theme_light()`, make BOTH X- and Y- axis have 15% larger font size, make the X-axis text blue, and make the Y-axis text purple.
```{r}
ggplot(iris, aes(x = Sepal.Length)) + 
  geom_histogram() + 
  theme_light() +
  theme(axis.text = element_text(size = rel(1.15)), # both axes 
        axis.text.x = element_text(color = "blue"), # only the x!
        axis.text.y = element_text(color = "purple")) # only the y!
```

<br>
You can **REMOVE** theme elements entirely with `element_blank()`. The example below is obviously bad plot design, but shows how to get the job done if you wanted to!
```{r}
ggplot(iris, aes(x = Sepal.Length)) + 
  geom_histogram() + 
  theme(axis.text = element_blank()) 
```

<br>
Rotate the text by a 90% angle
```{r}
ggplot(iris, aes(x = Sepal.Length)) + 
  geom_histogram() + 
  theme(axis.text = element_text(angle = 90))
```


### Quick examples of `element_rect()`

Change the background fill of the plot:
```{r}
ggplot(iris, aes(x = Sepal.Length)) + 
  geom_histogram() + 
  theme(plot.background = element_rect(fill = "pink"))
```

<br>
Change the background fill of the plot *PANEL* with `panel.backgroud`:
```{r}
ggplot(iris, aes(x = Sepal.Length)) + 
  geom_histogram() + 
  theme(panel.background = element_rect(fill = "pink"))
```

<br>
Change the background fill and color (and use `size` to increase line thickness) of the *facet (panel) strips*:
```{r, fig.width = 6}
ggplot(iris, aes(x = Sepal.Length)) + 
  geom_histogram() + 
  # Add faceting
  facet_wrap(vars(Species), nrow = 1) +
  theme(strip.background = element_rect(fill = "pink", 
                                        color = "blue",
                                        size = rel(4)))
```


### Quick examples of `element_line()`

Change the axis thickness (multiply default size by 2) and color, why not! 
```{r}
ggplot(iris, aes(x = Sepal.Length)) + 
  geom_histogram() + 
  theme(axis.line = element_line(size = rel(2), color = "purple"))
```

<br>
Change the background grid to blue! 
```{r}
ggplot(iris, aes(x = Sepal.Length)) + 
  geom_histogram() + 
  theme(panel.grid = element_line(color = "blue"))
```

<br>
Remove the *minor grid lines* (still blue to help you see the difference)
```{r}
ggplot(iris, aes(x = Sepal.Length)) + 
  geom_histogram() + 
  theme(panel.grid = element_line(color = "blue"),
        panel.grid.minor = element_blank())
```

## Modifying the theme of an existing plot


You can continue to add features, including theme changes, to a plot once it is created. For example:

```{r}
ggplot(iris, aes(x = Sepal.Length)) + 
  geom_histogram() +
  labs(title = "My iris histogram") -> iris_histogram # save plot to variable

# Add some theme changes to it, without changing the definition of `iris_histogram`
iris_histogram + 
  theme(plot.title = element_text(hjust = 0.5)) # moves title to top-MIDDLE of the plot instead of the default top-left
```

## Setting a theme for your entire session/script

Rather than adding a theme style onto each individual plot, you can use the function `theme_set()` set the theme to use for *all plots made in a given R session*. You can also include command in a script so that all plots made *on lines written AFTER* that command will automatically use that theme. 

+ To set a complete theme (for example, `theme_classic()` as the default, use: `theme_set(theme_classic())`
+ To set a customized theme, use: `theme_set(theme_classic() + theme(STUFF THAT NORMALLY GOES INTO THEME))`
