---
title: "Homework #1: Due Wednesday 9/15/21 to Canvas as a PDF by 2:00 pm"
author: "Data Science for Biologists, Fall 2021"
output: 
  html_document:
    theme: lumen
    highlight: tango
---
<style>
.main-container {
    max-width: 800px;
    margin-left: 100px;
    margin-right: auto;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(htmlwidgets.TOJSON_ARGS = list(na = 'string'))
library(DT)
library(readr)
library(dplyr)
seals <- read_csv("seals.csv")
set.seed(012021) # 01 for fall, 02 for spring
readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-07/coffee_ratings.csv') %>%
    dplyr::filter(category_two_defects > 0) %>%
    dplyr::select(total_cup_points,
                  species,
                  variety,
                  country_of_origin,
                  region,
                  farm_name,
                  number_of_bags,
                  number_of_defects = category_two_defects,
                  aroma,
                  flavor,
                  acidity,
                  balance,
                  sweetness,
                  moisture) %>%
    group_by(species) %>%
    sample_n(10) %>% 
    ungroup() -> coffee_ratings 


coffee_ratings <- coffee_ratings[sample(20),]
```

```{css, echo=F}
blockquote{
  margin-left: 25px;
  font-size: 1em;
}
```

## Homework Dataset Background

This dataset, called `coffee_ratings`, provides ratings of different coffee beans made around the world. Each row is a single coffee bean, and each column provides different information about that coffee bean including its total rating (variable `total_cup_points`, on a scale of 100), what species and variety of coffee it is, information about where the coffee comes from, information about the number of defects detected per coffee, and different graded measurements, on a scale of 10, for various coffee attributes (like aroma or sweetness). It has been subsetted (to only 20 rows and 14 columns) from the full dataset [here](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-07-07/readme.md).

You will see some entries in the dataset that are `NA`, which implies a *missing value*. Sometimes, data is missing or unknown, and it doesn't necessarily imply anything is wrong with the dataset. Instead, it implies that data collection can be challenging, and you can't always get all the information you want. 



### Explore the dataset 

> Use the "Show [] entries" dropdown menu below to see all 20 rows.
>
> Scroll to the right to see all columns.
>
> You can use the "Search" field to find values or variables in the data. _CAUTION: Does not work to search for `NA`s_.
> 
> Click on a column header to sort the data by that column.

```{r}
datatable(coffee_ratings, 
          rownames=F, 
          options = list(
            initComplete = JS(
              "function(settings, json) {",
              "$(this.api().table().header()).css({'background-color': '#996633', 'color': '#000'});",
              "}")
))
```


## Specific Instructions:


Answer these questions about the dataset variables and how one might visualize the data. **When referring to variable names in the data, you must be _absolutely precise with spelling and capitalization._** 

For example, there is *no variable* called `Total Cup Points` but there *is* a variable called `total_cup_points`.

You do not need to write in complete sentences _except_ for one of your answers in Set 2 (see specific instructions below).


### Set 1 

**Instructions: For these questions, simply list the variables.**

1. Which variable(s), if any, in this dataset is/are *categorical*?

2. Which variable(s), if any, in this dataset is/are *numeric continuous*?

3. Which variable(s), if any, in this dataset is/are *numeric discrete*?

4. For which variable(s), if any, do we **NOT** know all of the values? (in other words, there is missing data)?

<br><br>

### Set 2

**Instructions: For these questions, indicate if the presented information is a _variable_ or a _value_ in the dataset.** If it is a variable, provide one of the values associated with the variable (*don't make it up - find a value in the data!*). If it is a value, tell me what variable the value belongs to. 

For all items below, you CAN answer the first part (variable or value?). But, for ONE AND ONLY ONE of the items below, you can't definitively answer the second part of the question (if variable give me a value, or if value give me the variable). For this one _ambiguous_ item, explain in 1 sentence why it is ambiguous.


> As an example, `total_cup_points` is a variable, and one of its values is 83.3. By contrast, "Arabica" is a value that is associated with the variable `species`.
>
> Hint: Use the search bar above the table to help you find values!


1. `variety`

2. `India`

3. `1`

4. `Caturra`

5. `veracruz`

6. `sweetness`

7. `0.12`

8. `83.25`

9. `farm_name`

<br><br>

### Set 3

**Instructions: For these questions, simply answer whether the choice of data visualization is appropriate or not with "yes" or "no".** Your answers should be based ONLY on whether the type of plot is technically appropriate for visualizing those data types (don't worry about whether it might "look pretty" or not).

1. You make a *scatterplot* to visualize the distribution of `sweetness` coffee rating values.

2. You make a *barplot* to visualize how many coffees come from each `country_of_origin` in the dataset.

3. You make a *histogram* to visualize the distribution of `flavor` rating values.

4. You make a *violin plot* to visualize the distribution of `acidity` rating values for each `species` of coffee.

5. You make a *scatterplot* to visualize the relationship between coffee `balance` and `sweetness` rating values.

6. You make a *boxplot* to visualize how many coffees of each `species` there are. 

7. You make a *density plot* to visualize the relationship between the `number_of_bags` produced and the `number_of_defects` found for each coffee.

8. You make a *strip/jitter plot* to visualize the full distributions of `flavor` ratings for coffee grown in each `region`.

9. You make a *barplot* to visualize the full distribution of `total_cup_points` coffee ratings.

<br><br>

### Set 4

These questions consider _five different datasets_ that are not `coffee_ratings`, as shown below. For dataset, indicate if it is tidy or not by answering simply **yes or no.** Then...

+ If the data _is_ tidy, _state_ whether each variable is categorical, numeric continuous, or numeric discrete.
+ If the data is _not_ tidy, _state_ which of the three qualities of tidy data are violated (one or more can be violated at a time!). You can simply use **1**, **2**, and/or **3**, as described below...

Recall that tidy data has _all of_ these qualities:

1. All columns are _variables_
    + This means for each value in a column, you should be able to say "this value can be one of <column name>. For example, considering `coffee_ratings`, "Arabica" can be one of `species`.
2. All rows contain only a _single observation_
    + Each row should contain ONE measured data point. For example, each row in `coffee_ratings` contains information about ONE recorded coffee bean. 
    + _Hint_: This means that, if the data records an experimental result with multiple measured _results_ for a given individual, those results should be on separate rows.
3. Each value is a _single_ value; multiple values should not be stored in a single table cell. 

<br>


#### **Dataset 1**
This dataset shows results from an experiment assessing the effect of different pesticides (here, called `A`, `B`, ...`F`) on insect control in agriculture. Different plots were sprayed with each pesticide, and the number of insects in each plot were counted after a certain period of time.

```{r}
datatable(InsectSprays %>% rename(n_insects = count, pesticide = spray), 
          rownames=F, 
          options = list(columnDefs = list(list(className = 'dt-left', targets = '_all'))), width = "65%") %>%
  formatStyle(columns = 1:ncol(InsectSprays), `text-align` = 'left')
```

<br>


#### **Dataset 2**
This dataset shows results from an experiment assessing oxygen use in seals when they dive. Researchers were testing the hypothesis that seals use more oxygen during feeding dives (i.e., dive to hunt and eat) compared to non-feeding dives (i.e., a dive without hunting/eating). Oxygen usage was measured twice for each seal: Once on a non-feeding dive on once on a feeding dive.

```{r}
datatable(seals %>%
            tidyr::pivot_wider(names_from = feeding, values_from = oxygen_use) %>%
            rename(seal_number = seal, 
                   non_feeding_oxygen_usage = No,
                   feeding_oxygen_usage = Yes), rownames=F,
          options = list(columnDefs = list(list(className = 'dt-left', targets = '_all'))), width = "65%") %>%
  formatStyle(columns = 1:3, `text-align` = 'left')
```


<br>

#### **Dataset 3**
This dataset shows partial results from a case-control study of esophageal cancer which enrolled individuals with esophageal cancer ("cases") and individuals without cancer ("controls"). Researchers obtained data from each patient including their age, and how much alcohol and tobacco they consume per day (don't worry about units). This data shows the how many patients in each combination of age, alcohol usage, and tobacco usage groups were enrolled.

```{r}
datatable(esoph %>%
            rename(age_group = agegp,
                   alcohol_consumption = alcgp,
                   tobacco_consumption = tobgp,
                   n_patients_cases = ncases,
                   n_patients_controls = ncontrols) %>%
            mutate(alcohol_consumption = stringr::str_replace(alcohol_consumption, "g/day", ""),
                   tobacco_consumption = stringr::str_replace(alcohol_consumption, "g/day", "")), 
          rownames=F,
          options = list(columnDefs = list(list(className = 'dt-left', targets = '_all'))), width = "65%") %>%
  formatStyle(columns = 1:5, `text-align` = 'left')
```

<br>

#### **Dataset 4**
This dataset shows the number of monthly deaths from 1974-1979 from lung disease, including bronchitis, emphysema and asthma, in the UK.  

```{r}
ldeaths_tibble <- tribble(
~year,~Jan,  ~Feb,  ~Mar,  ~Apr,  ~May,  ~Jun,  ~Jul,  ~Aug,  ~Sep,  ~Oct,  ~Nov,  ~Dec,
1974, 3035, 2552, 2704, 2554, 2014, 1655, 1721, 1524, 1596, 2074, 2199, 2512,
1975, 2933, 2889, 2938, 2497, 1870, 1726, 1607, 1545, 1396, 1787, 2076, 2837,
1976, 2787, 3891, 3179, 2011, 1636, 1580, 1489, 1300, 1356, 1653, 2013, 2823,
1977, 3102, 2294, 2385, 2444, 1748, 1554, 1498, 1361, 1346, 1564, 1640, 2293,
1978, 2815, 3137, 2679, 1969, 1870, 1633, 1529, 1366, 1357, 1570, 1535, 2491,
1979, 3084, 2605, 2573, 2143, 1693, 1504, 1461, 1354, 1333, 1492, 1781, 1915
)
datatable(ldeaths_tibble, rownames=F,
          options = list(columnDefs = list(list(className = 'dt-left', targets = '_all'))), width = "65%") %>%
  formatStyle(columns = 1:13, `text-align` = 'left')
```

<br>



#### **Dataset 5**
This dataset shows the number of tuberculosis cases out of the total population in three different countries as recorded in years 1999 and 2000. 

```{r}
datatable(tidyr::table3 %>%
            mutate(rate = stringr::str_replace(rate, "/", " cases in total population of ")) %>%
            rename(cases_per_population = rate), rownames=F,
          options = list(columnDefs = list(list(className = 'dt-left', targets = '_all'))), width = "65%") %>%
  formatStyle(columns = 1:3, `text-align` = 'left')
```

<br>