---
title: "Tidy Tuesday Office"
author: "Poonam Shah"
date: "3/21/2020"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(schrute)
library(tidylog)
library(tidytext)
theme_set(theme_minimal())
```


```{r}
# Get the Office rating data
office_ratings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/
                                  tidytuesday/master/data/2020/2020-03-17/
                                  office_ratings.csv')
#Get the Schrute data
mydata <- schrute::theoffice
#convert a dataframe with a text column to be one-token-per-row
mydata %>%
  tidytext::unnest_tokens(words,text) ->Schrute_tidytext
```


```{r }
#Kelly the amount of words per season 
Schrute_tidytext %>%
  #filter the character by Kelly
  filter(character=="Kelly") %>%
  #deleted the unused columns
  select(-index,-episode,-episode_name,-director,
         -writer,-imdb_rating,-air_date,-total_votes, -text_w_direction)%>%
  group_by(season,character) %>%
  count() -> Mindy_Project

#Making the plot
Mindy_Project %>%
  ggplot(aes(x=season, y=n, fill=season)) +
  geom_bar(stat = "identity") +
  #Adding a title and cleaning up the axis
  labs(x="Season", y="Number of Words", 
       title="Kelly the amount of words per season") +
  scale_color_brewer(palette = "Set1") -> Mindy_Project_plot

print(Mindy_Project_plot)

#Saving the plot as a picture
ggsave("Mindy_Project_plot.png", Mindy_Project_plot, height = 4, width = 6)

```

This week TidyTuesday was using a new package called tidytext. It is a package to help in text mining dataset and books. I used with package to see how amount words the character Kelly said in every season. The tidytext package was great in The Office dataset. I have only scratched the surface of the tidytext package I would love to practice more in the future for other tidytuesday.
