---
title: 'Tidy Tuesday Week 1 (week 8 in regards to the data)'
author: "Paige Richards"
date: "2/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gganimate)
library(tidytuesdayR)
library(patchwork)
```


```{r}
tidytuesdayR::tt_load(2020, week=8) -> tidytuesdata
view(tidytuesdata)
tidytuesdata$food_consumption -> tidytuesdata2
print(tidytuesdata2)
```

```{r, fig.height=6,fig.width=10}
tidytuesdata2 %>%
  dplyr::filter(country==c('Kazakhstan','Egypt'))%>%
  dplyr::filter(food_category=='Eggs')%>%
  ggplot(aes(y=co2_emmission, x=country))+
    geom_col(aes(fill=country), width=0.5) +
    scale_fill_manual(values=c('darkorange', 'darkred'))+
  theme_classic()+
  labs(y='CO2 Emmission', x='Country')+
  theme(legend.position='none')+
 theme(plot.title = element_text(size = rel(1.8)), axis.title=element_text(size=rel(1)), axis.text= element_text(size=rel(1)), axis.ticks.length.x= unit(0.3, "cm"), axis.ticks.length.y=unit(0.3, 'cm'))+
  ggtitle('Consumption of Eggs')-> eggs_graph
tidytuesdata2 %>%
  dplyr::filter(country==c('Kazakhstan','Egypt'))%>%
  dplyr::filter(food_category=='Soybeans')%>%
  ggplot(aes(y=co2_emmission, x=country))+
    geom_col(aes(fill=country), width=0.5)+
  scale_fill_manual(values=c('darkorange', 'darkred'))+
  theme_classic()+
  labs(y='CO2 Emmission', x='Country')+
  theme(legend.position='none')+ 
  theme(plot.title = element_text(size = rel(1.8)), axis.title=element_text(size=rel(1)),axis.text= element_text(size=rel(1)), axis.ticks.length.x= unit(0.3, "cm"), axis.ticks.length.y=unit(0.3, 'cm'))+
  ggtitle('Consumption of Soybeans')-> soybeans_graph
soybeans_graph + eggs_graph -> soybean_egg_combined_graph
print(soybean_egg_combined_graph)
ggsave('tidy_tuesday_graph.png',soybean_egg_combined_graph)
```

Although this plot has a **ton** of code that is unnecessary and able to be distilled down, it was extremely helpful to write it this way. I tried to use gganimate, but I found that it may not be a useful technique for this type of information. I had a graph that was slightly animated but it looked confusing and did not make sense for the information I was trying to convey. I found that it may be best used for something that has time included in the data. I gained experience and practice with dplyr and patchwork! I feel more comfortable understanding how dplyr filtering works and creating a graph from it. It was also helpful learning how to wrangle this data to figure out what I wanted to show. I learned how to make the text larger of the title, axis, and axis ticks! It was interesting to learn all of the ways I could customize the graph.