---
title: Launching "Data Science for Biologists" exercises
output: 
  html_document:
    theme: lumen
    highlight: tango
---



## Apps
<br>

#### "Types of plots"

+ [Access via browser, _on-campus_](http://10.33.168.74:3838/spielman/types.of.plots)
+ [Access via browser, _off-campus_](https://sjspielman.shinyapps.io/types_of_plots)
+ Access via the `ds4b.materials` library in RStudio Cloud:

  ```{r eval=F}
  library(ds4b.materials)
  launch_app("types_of_plots")
  ```


#### "Line by line"

+ [Access via browser, _on-campus_](http://10.33.168.74:3838/spielman/ds4b.materials/inst/apps/line_by_line/)
+ [Access via browser, _off-campus_](https://sjspielman.shinyapps.io/line_by_line)
+ Access via the `ds4b.materials` library in RStudio Cloud:

  ```{r eval=F}
  library(ds4b.materials)
  launch_app("line_by_line")
  ```

<br><br>


## Exercises 

<br>

#### "Types of data and figures"

+ [Access via browser _on-campus_](http://10.33.168.74:3838/spielman/ds4b.materials/inst/tutorials/data_figures/)
+ Access via the `ds4b.materials` library in RStudio Cloud:

  ```{r eval=F}
  library(ds4b.materials)
  launch_app("data_figures")
  ```

#### "Introduction to R"

+ [Access via browser _on-campus_](http://10.33.168.74:3838/spielman/ds4b.materials/inst/tutorials/intro_R)
+ Access via the `ds4b.materials` library in RStudio Cloud:

  ```{r eval=F}
  library(ds4b.materials)
  launch_app("intro_R")
  ```


#### "Introduction to `ggplot2`"

+ [Access via browser _on-campus_](http://10.33.168.74:3838/spielman/ds4b.materials/inst/tutorials/ggplot2)
+ Access via the `ds4b.materials` library in RStudio Cloud:

  ```{r eval=F}
  library(ds4b.materials)
  launch_app("ggplot2")
  ```


#### "Introduction to `dplyr`"

+ [Access via browser _on-campus_](http://10.33.168.74:3838/spielman/ds4b.materials/inst/tutorials/dplyr)
+ Access via the `ds4b.materials` library in RStudio Cloud:

  ```{r eval=F}
  library(ds4b.materials)
  launch_app("dplyr")
  ```
