---
title: 'Instructions: Homework #3'
author: "Data Science for Biologists, Fall 2021"
date: "Complete the template Rmd and submit to Canvas on 9/21/21 by 1 PM"
output: 
  html_document:
    theme: lumen
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

> For this homework, you will be writing your very first _RMarkdown file_ and practicing your base R skills. 

## Instructions

+ Obtain the homework from your RStudio Cloud class project by running the following code in the R Console:

  ```{r, eval=FALSE}
  library(ds4b.materials) # Load the class library
  launch_homework(3)      # Launch Homework 3
  ```

+ Answer each question with appropriate code and comments in the given question's R chunk. Chunks are _named_ as `q1` for question 1, `q2` for question 2, etc.
  + Always ensure your answers are being produced by code. No credit will be given for answers that are not directly produced by code. Please refer to the [Homework #2 Instructions](../hw2/hw2_instructions.html) for examples of acceptable and unacceptable code.
  + You should have one comment per line of code. Comments should be written in "plain english" and give _brief descriptions_ about what the code is doing. Comments should be written _above_ the line of code they refer to. You will not be graded on whether I think your comment is "good enough" - your comments need only be "good enough" for YOU to understand what your code is doing.
<br><br>
+ You must set an **RMarkdown theme and code syntax highlighting scheme** of your choosing in the YAML front matter. These links will help you:
  + Set the theme and syntax highlighting according to this demonstration video 
  + Choose your favorite _theme_ among the **pre-packaged themes** (ignore everything below "Even More Themes") shown at [this link](https://www.datadreaming.org/post/r-markdown-theme-gallery/)
  + Choose your favorite _syntax highlighting_ among these options at [this link](https://www.garrickadenbuie.com/blog/pandoc-syntax-highlighting-examples/)
<br><br>
+ Make sure to use the `ds4b.materials` function `check_homework(3)` to **check** whether there are any bugs in your homework. Make sure to find and fix all bugs before submitting! Your document _must knit without errors_ for full credit, so don't skip this step!
<br><br>
+ Additional tips for success:
  + _**Start the homework EARLY. Do not wait until the day before the deadline to begin.**_
  + Remember to use the `introverse` library interactively in the R Console to get help with the function `get_help()`.
  + Remember that R code chunks behave interactively so you do not need to use the `print()` function to print. Simply type a variable's name to reveal its contents, just like in the Console.


## Homework Questions


1. Define two variables `a` and `b` as values 8 and 64, respectively. Use a logical operator and the function `sqrt()` to ask if `a` is equal to the square root of `b`, and print the output. **The code should print `TRUE` if they are equal and `FALSE` if not.**
    + Need help? Run `get_help("logical")` and `get_help("sqrt")` in the R Console after loading the `introverse` library.

<br> 

2. Define a variable `success_status` whose values *depends* on a logical comparison, using variables `a` and `b` from the previous question. If `a` squared equals `b`, `success_status` should be defined with the value "eureka". Otherwise, if `a` squared does not equal `b`, `success_status` should be defined with the value "wompwomp". **Print the value of `success_status` after it is defined.**
      + Need help? Run `get_help("ifelse")` in the R Console after loading the `introverse` library.
      + Do NOT redefine the `a` and `b` variables. You defined them in question #2, and you are still coding in the same Rmarkdown file. _We always want to avoid repeating code since it often leads to unexpected bugs._

<br> 

3. Define an array variable called `mammal_vectors` to contain these strings: "camel", "civet", "mink", "bat", "ferret", and "pangolin" (these are all mammals known to be vectors for either SARS-CoV-2 or one of its close relatives SARS and MERS).  Use the logical operator `%in%` to determine if the string "elephant" is in the array. **If "elephant" is in the array, your code should print `TRUE`, and `FALSE` otherwise.**
    + Need help? Run `get_help("logical")` in the R Console after loading the `introverse` library.

<br>

4. Define an array variable called `dinosaurs` to contain the three strings: "eagle", "ostrich", and "sparrow" (wait, those are all birds?! That's right, birds are dinosaurs!). Then, define two other variables called `mammals_length` and `dinosaurs_length`, which should each contain the result of running `length()` on each array. Finally, use a logical operator to compare these variables. **If `mammals_length` is larger than `dinosaurs_length` (in other words, the mammals array is longer than the dinosaurs array), your code should print `TRUE`. Otherwise, your code should print `FALSE`.**
    + Need help? Run `get_help("length")` in the R Console after loading the `introverse` library.

<br> 

5. Use the `nchar()` function to determine how many characters are in each string in the `dinosaurs` array, and save the result of this operation to a variable `nchar_dinosaurs`. Then, use a logical operator to ask if each value is equal to 7. **Your code should print a new array containing `TRUE` or `FALSE` values representing whether this condition was met for each value.**
    + Need help? Run `get_help("nchar")` in the R Console after loading the `introverse` library.


<br> 

6. Use a colon (`:`) to create a variable called `q6` as an array of numbers in order -12 through 17. From this variable `q6`, calculate the mean, median, standard deviation, and sum (with functions `mean()`, `median()`, `sd()`,and `sum()`, respectively) of this array. Each calculation should be saved to a variable called, respectively: `q6_mean`, `q6_median`, `q6_sd`, and `q6_sum`. **Finally, print these four variables. You do not need to print `q6`.**
    + Need help? Run `get_help("mean")` in the R Console after loading the `introverse` library.

<br> 

7. Again consider the array `q6`. Use a logical operator to ask if all values in this array are less than or 0 (aka, is `q6` less than or equal to 0?). The result of this code will be an array itself containing `TRUE` and `FALSE` values corresponding to whether the $\leq0$ condition was met for each value in `q6`.  Save this resulting array to a variable called `q6_leq0` ("leq" = "**l**ess than or **eq**ual to"), and **print the value of `q6_leq0`.**
    + Need help? Run `get_help("logical")` in the R Console after loading the `introverse` library.

<br> 


8. Teach yourself how to use the function `all()` using the `introverse` help docs, by running `get_help("all")` in the R Console, after you've loaded the `introverse` library. Then, use this function to determine whether all values of `q6` are _less than or equal to 0._ **Your code should print a _single_ `TRUE` if all values are less than or equal to 0, and a single `FALSE` otherwise.**
    + Hint: The code is _almost exactly identical_ to the code for question 7!
    
<br> 

9. Again consider the array `q6`. For this question, find the _sum_ of the _absolute values_ of this array. To achieve this, you will need to first get the absolute value of the array `q6` (using the `abs()` function), and then you will need to sum those values (using the `sum()` function). **Your code chunk should reveal the final sum.**
    + Need help? Run `get_help("sum")` or `get_help("abs")` in the R Console after loading the `introverse` library.
    + Hint: You can approach this question in several ways, as long as you use `abs()` and then `sum()` on `q6` in that order.
  

<br> 


10. There is a super useful function called `paste()` which allows you to strings (characters) together to create a new string. See how this function works below:

      ```{r paste_demo}
      combined_string <- paste("string 1", "and", "string 2")
      print(combined_string)
      
      # you can also include variables:
      a_string <- "hello there"
      combined_string2 <- paste("string 1", a_string, "string 2")
      print(combined_string2)
      ```

    **Use the `paste()` function and the variable `q6_mean` you previously defined to print a sentence that says: "The mean of q6 is ...." where .... is the the `q6_mean` value.** Your code should _NOT_ include a number directly (don't copy/paste the value of the mean). You may _ONLY_ refer to this number value with the `q6_mean` variable.
        + Need help? Run `get_help("paste")` in the R Console after loading the `introverse` library.

<br>


11. For the remaining questions, you will consider the built-in dataset `iris`, a famous dataset with 150 rows and 5 columns that provides physical measurements for 150 iris specimens from three different species.

    For this question, write _code_ to determine the following information about `iris`. You need to use a single function for each task (there will be four lines of code in the end), and all output should be printed from the chunk.

    + The number of rows
    + The number of columns
    + The _names_ of the columns (use a function that here that only tells you the names!)
    + A _summary_ of the distribution of each column in the data


<br>

12. The goal of this question is to answer, "do sepal lengths or petal lengths have more variation?" When comparing variation across _different variables_, we can calculate the _coefficient of variation_ (COV) which is the standard deviation divided by the mean. Calculate the COV for sepal lengths and petal lengths, and save to variables `cov_sepal_length` and `cov_petal_length`, respectively. Then, use a logical operator to ask whether sepal lengths have _more_ (`>`) variation than petal lengths. **Your code should print `TRUE` if sepal lengths have more spread, and `FALSE` otherwise.**

<br>

13. Calculate the mean petal width and save the result to a variable `mean_petal_width`. Then, use `paste()` to print a sentence that reads: `"The mean petal width in the iris dataset is ..."`, where `...` is replaced with the value in `mean_petal_width`.

<br>

14. The `iris` variable `Species` is a _factor variable_, meaning R recognizes it as containing _categories_. By contrast, the variable `Sepal.Width` is numeric, not a factor, so R doesn't recognize that it has categories. Use the `levels()` function separately on each of these functions to determine the different categories in the column. **Your code should simply print the result of running the `levels()` function on those columns.**
    + Hint: The code output should be consistent with whether the variables are factors. Non-factor variables don't have any levels and the code result will reflect this.

<br>

15. To learn a little more about factors, let's _coerce_ some columns into and away from factors. To do this, we'll create a new dataset called `iris2` containing some coerced columns. Copy and paste the following code into the R chunk for this question (this code is part of the answer!). This code creates a new data frame `iris2` that contains the same information as `iris`, but `Species` is no longer factor (it is now character), and `Sepal.Width` is no longer numeric (it is now a factor).

    ```{r eval=F}
    # Create a copy of iris in iris2
    iris2 <- iris
    
    # Coerce Species into a character
    iris2$Species <- as.character(iris2$Species)
    
    # Coerce Sepal.Width into a factor
    iris2$Sepal.Width <- as.factor(iris2$Sepal.Width)
    ```
    
    Again write code with `levels()` to see the categories in the **`iris2`** columns `Species` and `Sepal.Width`, and notice how the output differs. Finally, write two more line of code in this chunk to **calculate the mean of the `Sepal.Width` column in `iris` AND in `iris2`**, which should also be printed out. In ~1 sentence of written text _below the code chunk_, explain why _IN YOUR OWN WORDS_ this calculation didn't work as well as you might have hoped (hint: the output from `levels()` is informative!). 